### 2018-06-30 figures/Hbb/fit/base_ebv2_signal_cH_default_unblinded/Tables/Yields_postFit.tex

\documentclass[10pt]{article}
\usepackage{siunitx}
\usepackage[margin=0.1in,landscape,papersize={210mm,350mm}]{geometry}
\begin{document}
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline 
 & 4jex-2bex & 4jex-3bex & 4jex-4bin & 5jex-2bex & 5jex-3bex & 5jex-4bin & 6jin-2bex & 6jin-3bex & 6jin-4bin\\
\hline 
  t#bar{t}+light   & \num[round-mode=figures,round-precision=3]{82731.6} $\pm$ \num[round-mode=figures,round-precision=3]{3743.39} & \num[round-mode=figures,round-precision=3]{4864.82} $\pm$ \num[round-mode=figures,round-precision=3]{443.353} & \num[round-mode=figures,round-precision=3]{15.3037} $\pm$ \num[round-mode=figures,round-precision=3]{11.7737} & \num[round-mode=figures,round-precision=3]{37928.2} $\pm$ \num[round-mode=figures,round-precision=3]{2705} & \num[round-mode=figures,round-precision=3]{3468.45} $\pm$ \num[round-mode=figures,round-precision=3]{385.694} & \num[round-mode=figures,round-precision=3]{15.7506} $\pm$ \num[round-mode=figures,round-precision=3]{9.31776} & \num[round-mode=figures,round-precision=3]{20031.7} $\pm$ \num[round-mode=figures,round-precision=3]{2207.33} & \num[round-mode=figures,round-precision=3]{2547.58} $\pm$ \num[round-mode=figures,round-precision=3]{415.229} & \num[round-mode=figures,round-precision=3]{20.7941} $\pm$ \num[round-mode=figures,round-precision=3]{18.6547} \\ 
  t#bar{t}+cc   & \num[round-mode=figures,round-precision=3]{11413.6} $\pm$ \num[round-mode=figures,round-precision=3]{4745.82} & \num[round-mode=figures,round-precision=3]{1354.89} $\pm$ \num[round-mode=figures,round-precision=3]{558.336} & \num[round-mode=figures,round-precision=3]{5.86827} $\pm$ \num[round-mode=figures,round-precision=3]{4.11578} & \num[round-mode=figures,round-precision=3]{8341.88} $\pm$ \num[round-mode=figures,round-precision=3]{3373.8} & \num[round-mode=figures,round-precision=3]{1996.21} $\pm$ \num[round-mode=figures,round-precision=3]{769.437} & \num[round-mode=figures,round-precision=3]{39.2465} $\pm$ \num[round-mode=figures,round-precision=3]{18.1406} & \num[round-mode=figures,round-precision=3]{7794.07} $\pm$ \num[round-mode=figures,round-precision=3]{3184.81} & \num[round-mode=figures,round-precision=3]{2975.91} $\pm$ \num[round-mode=figures,round-precision=3]{1151.64} & \num[round-mode=figures,round-precision=3]{58.5315} $\pm$ \num[round-mode=figures,round-precision=3]{25.1281} \\ 
  t#bar{t}+bb   & \num[round-mode=figures,round-precision=3]{4274.97} $\pm$ \num[round-mode=figures,round-precision=3]{586.681} & \num[round-mode=figures,round-precision=3]{3403.3} $\pm$ \num[round-mode=figures,round-precision=3]{355.431} & \num[round-mode=figures,round-precision=3]{109.855} $\pm$ \num[round-mode=figures,round-precision=3]{16.9385} & \num[round-mode=figures,round-precision=3]{3410.91} $\pm$ \num[round-mode=figures,round-precision=3]{467.915} & \num[round-mode=figures,round-precision=3]{4920.81} $\pm$ \num[round-mode=figures,round-precision=3]{465.43} & \num[round-mode=figures,round-precision=3]{355.754} $\pm$ \num[round-mode=figures,round-precision=3]{28.8725} & \num[round-mode=figures,round-precision=3]{3395.32} $\pm$ \num[round-mode=figures,round-precision=3]{488.592} & \num[round-mode=figures,round-precision=3]{7518.29} $\pm$ \num[round-mode=figures,round-precision=3]{767.709} & \num[round-mode=figures,round-precision=3]{1105.48} $\pm$ \num[round-mode=figures,round-precision=3]{81.9048} \\ 
  topEW   & \num[round-mode=figures,round-precision=3]{173.141} $\pm$ \num[round-mode=figures,round-precision=3]{27.3722} & \num[round-mode=figures,round-precision=3]{34.8919} $\pm$ \num[round-mode=figures,round-precision=3]{5.71662} & \num[round-mode=figures,round-precision=3]{2.67875} $\pm$ \num[round-mode=figures,round-precision=3]{0.540039} & \num[round-mode=figures,round-precision=3]{167.465} $\pm$ \num[round-mode=figures,round-precision=3]{25.9124} & \num[round-mode=figures,round-precision=3]{65.3104} $\pm$ \num[round-mode=figures,round-precision=3]{10.4221} & \num[round-mode=figures,round-precision=3]{8.15336} $\pm$ \num[round-mode=figures,round-precision=3]{1.43885} & \num[round-mode=figures,round-precision=3]{212.764} $\pm$ \num[round-mode=figures,round-precision=3]{33.8202} & \num[round-mode=figures,round-precision=3]{146.887} $\pm$ \num[round-mode=figures,round-precision=3]{23.862} & \num[round-mode=figures,round-precision=3]{26.9491} $\pm$ \num[round-mode=figures,round-precision=3]{4.79349} \\ 
  ttH   & \num[round-mode=figures,round-precision=3]{62.4444} $\pm$ \num[round-mode=figures,round-precision=3]{7.53342} & \num[round-mode=figures,round-precision=3]{47.1158} $\pm$ \num[round-mode=figures,round-precision=3]{6.06513} & \num[round-mode=figures,round-precision=3]{4.65928} $\pm$ \num[round-mode=figures,round-precision=3]{0.676057} & \num[round-mode=figures,round-precision=3]{81.0323} $\pm$ \num[round-mode=figures,round-precision=3]{8.86474} & \num[round-mode=figures,round-precision=3]{98.5241} $\pm$ \num[round-mode=figures,round-precision=3]{11.579} & \num[round-mode=figures,round-precision=3]{16.5515} $\pm$ \num[round-mode=figures,round-precision=3]{2.31589} & \num[round-mode=figures,round-precision=3]{134.587} $\pm$ \num[round-mode=figures,round-precision=3]{15.1851} & \num[round-mode=figures,round-precision=3]{240.518} $\pm$ \num[round-mode=figures,round-precision=3]{29.1002} & \num[round-mode=figures,round-precision=3]{61.6042} $\pm$ \num[round-mode=figures,round-precision=3]{8.77664} \\ 
  W+jetslight   & \num[round-mode=figures,round-precision=3]{6.12582} $\pm$ \num[round-mode=figures,round-precision=3]{6.30201} & \num[round-mode=figures,round-precision=3]{0.0116465} $\pm$ \num[round-mode=figures,round-precision=3]{0.0131854} & \num[round-mode=figures,round-precision=3]{4.05447e-06} $\pm$ \num[round-mode=figures,round-precision=3]{4.98735e-06} & \num[round-mode=figures,round-precision=3]{2.43672} $\pm$ \num[round-mode=figures,round-precision=3]{2.37869} & \num[round-mode=figures,round-precision=3]{0.00827649} $\pm$ \num[round-mode=figures,round-precision=3]{0.00936765} & \num[round-mode=figures,round-precision=3]{5.90829e-06} $\pm$ \num[round-mode=figures,round-precision=3]{7.22092e-06} & \num[round-mode=figures,round-precision=3]{1.14449} $\pm$ \num[round-mode=figures,round-precision=3]{1.08589} & \num[round-mode=figures,round-precision=3]{0.00749858} $\pm$ \num[round-mode=figures,round-precision=3]{0.00862511} & \num[round-mode=figures,round-precision=3]{9.96564e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.24703e-05} \\ 
  W+jetscharm   & \num[round-mode=figures,round-precision=3]{115.305} $\pm$ \num[round-mode=figures,round-precision=3]{56.3834} & \num[round-mode=figures,round-precision=3]{1.21567} $\pm$ \num[round-mode=figures,round-precision=3]{0.666388} & \num[round-mode=figures,round-precision=3]{0.00240026} $\pm$ \num[round-mode=figures,round-precision=3]{0.00148325} & \num[round-mode=figures,round-precision=3]{47.0596} $\pm$ \num[round-mode=figures,round-precision=3]{22.8395} & \num[round-mode=figures,round-precision=3]{0.71657} $\pm$ \num[round-mode=figures,round-precision=3]{0.39251} & \num[round-mode=figures,round-precision=3]{0.00222823} $\pm$ \num[round-mode=figures,round-precision=3]{0.00141395} & \num[round-mode=figures,round-precision=3]{21.4321} $\pm$ \num[round-mode=figures,round-precision=3]{10.9827} & \num[round-mode=figures,round-precision=3]{0.608391} $\pm$ \num[round-mode=figures,round-precision=3]{0.342486} & \num[round-mode=figures,round-precision=3]{0.00339913} $\pm$ \num[round-mode=figures,round-precision=3]{0.00224384} \\ 
  W+jetsbeauty   & \num[round-mode=figures,round-precision=3]{4726.45} $\pm$ \num[round-mode=figures,round-precision=3]{1836.48} & \num[round-mode=figures,round-precision=3]{260.312} $\pm$ \num[round-mode=figures,round-precision=3]{104.409} & \num[round-mode=figures,round-precision=3]{2.88562} $\pm$ \num[round-mode=figures,round-precision=3]{1.346} & \num[round-mode=figures,round-precision=3]{2051.07} $\pm$ \num[round-mode=figures,round-precision=3]{819.003} & \num[round-mode=figures,round-precision=3]{168.887} $\pm$ \num[round-mode=figures,round-precision=3]{68.0043} & \num[round-mode=figures,round-precision=3]{6.0271} $\pm$ \num[round-mode=figures,round-precision=3]{2.82667} & \num[round-mode=figures,round-precision=3]{1185.98} $\pm$ \num[round-mode=figures,round-precision=3]{472.732} & \num[round-mode=figures,round-precision=3]{183.822} $\pm$ \num[round-mode=figures,round-precision=3]{75.4584} & \num[round-mode=figures,round-precision=3]{12.5462} $\pm$ \num[round-mode=figures,round-precision=3]{5.73829} \\ 
  Z+jetslight   & \num[round-mode=figures,round-precision=3]{1.61066} $\pm$ \num[round-mode=figures,round-precision=3]{1.54715} & \num[round-mode=figures,round-precision=3]{0.00375032} $\pm$ \num[round-mode=figures,round-precision=3]{0.00414649} & \num[round-mode=figures,round-precision=3]{1.39446e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.64694e-06} & \num[round-mode=figures,round-precision=3]{0.659446} $\pm$ \num[round-mode=figures,round-precision=3]{0.630272} & \num[round-mode=figures,round-precision=3]{0.00250007} $\pm$ \num[round-mode=figures,round-precision=3]{0.00279867} & \num[round-mode=figures,round-precision=3]{1.85341e-06} $\pm$ \num[round-mode=figures,round-precision=3]{2.26403e-06} & \num[round-mode=figures,round-precision=3]{0.302002} $\pm$ \num[round-mode=figures,round-precision=3]{0.273111} & \num[round-mode=figures,round-precision=3]{0.00216257} $\pm$ \num[round-mode=figures,round-precision=3]{0.00245868} & \num[round-mode=figures,round-precision=3]{2.97446e-06} $\pm$ \num[round-mode=figures,round-precision=3]{3.69548e-06} \\ 
  Z+jetscharm   & \num[round-mode=figures,round-precision=3]{20.2778} $\pm$ \num[round-mode=figures,round-precision=3]{9.77507} & \num[round-mode=figures,round-precision=3]{0.192583} $\pm$ \num[round-mode=figures,round-precision=3]{0.105433} & \num[round-mode=figures,round-precision=3]{0.000330274} $\pm$ \num[round-mode=figures,round-precision=3]{0.000208386} & \num[round-mode=figures,round-precision=3]{8.36981} $\pm$ \num[round-mode=figures,round-precision=3]{4.04635} & \num[round-mode=figures,round-precision=3]{0.127165} $\pm$ \num[round-mode=figures,round-precision=3]{0.0699958} & \num[round-mode=figures,round-precision=3]{0.000455404} $\pm$ \num[round-mode=figures,round-precision=3]{0.000281893} & \num[round-mode=figures,round-precision=3]{4.46059} $\pm$ \num[round-mode=figures,round-precision=3]{2.19994} & \num[round-mode=figures,round-precision=3]{0.124792} $\pm$ \num[round-mode=figures,round-precision=3]{0.0687308} & \num[round-mode=figures,round-precision=3]{0.000758631} $\pm$ \num[round-mode=figures,round-precision=3]{0.00047388} \\ 
  Z+jetsbeauty   & \num[round-mode=figures,round-precision=3]{1857.53} $\pm$ \num[round-mode=figures,round-precision=3]{727.411} & \num[round-mode=figures,round-precision=3]{102.596} $\pm$ \num[round-mode=figures,round-precision=3]{41.3086} & \num[round-mode=figures,round-precision=3]{1.9331} $\pm$ \num[round-mode=figures,round-precision=3]{1.03534} & \num[round-mode=figures,round-precision=3]{698.171} $\pm$ \num[round-mode=figures,round-precision=3]{274.747} & \num[round-mode=figures,round-precision=3]{74.3953} $\pm$ \num[round-mode=figures,round-precision=3]{30.343} & \num[round-mode=figures,round-precision=3]{5.58037} $\pm$ \num[round-mode=figures,round-precision=3]{3.22145} & \num[round-mode=figures,round-precision=3]{353.402} $\pm$ \num[round-mode=figures,round-precision=3]{146.179} & \num[round-mode=figures,round-precision=3]{56.5389} $\pm$ \num[round-mode=figures,round-precision=3]{23.7814} & \num[round-mode=figures,round-precision=3]{3.53814} $\pm$ \num[round-mode=figures,round-precision=3]{2.14836} \\ 
  Single-top   & \num[round-mode=figures,round-precision=3]{6394.03} $\pm$ \num[round-mode=figures,round-precision=3]{883.786} & \num[round-mode=figures,round-precision=3]{398.806} $\pm$ \num[round-mode=figures,round-precision=3]{70.2783} & \num[round-mode=figures,round-precision=3]{8.12021} $\pm$ \num[round-mode=figures,round-precision=3]{3.29235} & \num[round-mode=figures,round-precision=3]{2586.28} $\pm$ \num[round-mode=figures,round-precision=3]{545.244} & \num[round-mode=figures,round-precision=3]{325.205} $\pm$ \num[round-mode=figures,round-precision=3]{73.1068} & \num[round-mode=figures,round-precision=3]{13.5356} $\pm$ \num[round-mode=figures,round-precision=3]{4.21865} & \num[round-mode=figures,round-precision=3]{1247.31} $\pm$ \num[round-mode=figures,round-precision=3]{360.081} & \num[round-mode=figures,round-precision=3]{321.943} $\pm$ \num[round-mode=figures,round-precision=3]{98.5638} & \num[round-mode=figures,round-precision=3]{27.8892} $\pm$ \num[round-mode=figures,round-precision=3]{10.1643} \\ 
  Dibosons   & \num[round-mode=figures,round-precision=3]{242.502} $\pm$ \num[round-mode=figures,round-precision=3]{84.0284} & \num[round-mode=figures,round-precision=3]{16.2373} $\pm$ \num[round-mode=figures,round-precision=3]{5.71578} & \num[round-mode=figures,round-precision=3]{0.494856} $\pm$ \num[round-mode=figures,round-precision=3]{0.222674} & \num[round-mode=figures,round-precision=3]{111.655} $\pm$ \num[round-mode=figures,round-precision=3]{47.8353} & \num[round-mode=figures,round-precision=3]{12.47} $\pm$ \num[round-mode=figures,round-precision=3]{5.44917} & \num[round-mode=figures,round-precision=3]{0.764199} $\pm$ \num[round-mode=figures,round-precision=3]{0.392172} & \num[round-mode=figures,round-precision=3]{82.228} $\pm$ \num[round-mode=figures,round-precision=3]{40.0442} & \num[round-mode=figures,round-precision=3]{16.7013} $\pm$ \num[round-mode=figures,round-precision=3]{8.21106} & \num[round-mode=figures,round-precision=3]{1.70033} $\pm$ \num[round-mode=figures,round-precision=3]{0.901524} \\ 
  SM-4tops   & \num[round-mode=figures,round-precision=3]{0.102052} $\pm$ \num[round-mode=figures,round-precision=3]{0.0378717} & \num[round-mode=figures,round-precision=3]{0.0964359} $\pm$ \num[round-mode=figures,round-precision=3]{0.0364277} & \num[round-mode=figures,round-precision=3]{0.00739477} $\pm$ \num[round-mode=figures,round-precision=3]{0.00590938} & \num[round-mode=figures,round-precision=3]{0.367211} $\pm$ \num[round-mode=figures,round-precision=3]{0.126578} & \num[round-mode=figures,round-precision=3]{0.491049} $\pm$ \num[round-mode=figures,round-precision=3]{0.156563} & \num[round-mode=figures,round-precision=3]{0.0540613} $\pm$ \num[round-mode=figures,round-precision=3]{0.0203091} & \num[round-mode=figures,round-precision=3]{5.14473} $\pm$ \num[round-mode=figures,round-precision=3]{1.54452} & \num[round-mode=figures,round-precision=3]{19.1387} $\pm$ \num[round-mode=figures,round-precision=3]{5.77773} & \num[round-mode=figures,round-precision=3]{4.96117} $\pm$ \num[round-mode=figures,round-precision=3]{1.54499} \\ 
  QCDE   & \num[round-mode=figures,round-precision=3]{7789.54} $\pm$ \num[round-mode=figures,round-precision=3]{3437.44} & \num[round-mode=figures,round-precision=3]{600.731} $\pm$ \num[round-mode=figures,round-precision=3]{229.327} & \num[round-mode=figures,round-precision=3]{20.0736} $\pm$ \num[round-mode=figures,round-precision=3]{14.1015} & \num[round-mode=figures,round-precision=3]{2657.64} $\pm$ \num[round-mode=figures,round-precision=3]{935.089} & \num[round-mode=figures,round-precision=3]{486.701} $\pm$ \num[round-mode=figures,round-precision=3]{221.708} & \num[round-mode=figures,round-precision=3]{3.606} $\pm$ \num[round-mode=figures,round-precision=3]{4.81275} & \num[round-mode=figures,round-precision=3]{1220.05} $\pm$ \num[round-mode=figures,round-precision=3]{500.496} & \num[round-mode=figures,round-precision=3]{747.705} $\pm$ \num[round-mode=figures,round-precision=3]{336.918} & \num[round-mode=figures,round-precision=3]{28.4241} $\pm$ \num[round-mode=figures,round-precision=3]{24.6852} \\ 
  QCDMU   & \num[round-mode=figures,round-precision=3]{1250.71} $\pm$ \num[round-mode=figures,round-precision=3]{637.813} & \num[round-mode=figures,round-precision=3]{209.581} $\pm$ \num[round-mode=figures,round-precision=3]{98.6358} & \num[round-mode=figures,round-precision=3]{9.02057} $\pm$ \num[round-mode=figures,round-precision=3]{8.46944} & \num[round-mode=figures,round-precision=3]{737.502} $\pm$ \num[round-mode=figures,round-precision=3]{369.549} & \num[round-mode=figures,round-precision=3]{71.1185} $\pm$ \num[round-mode=figures,round-precision=3]{44.7342} & \num[round-mode=figures,round-precision=3]{2e-05} $\pm$ \num[round-mode=figures,round-precision=3]{1.00366e-05} & \num[round-mode=figures,round-precision=3]{307.593} $\pm$ \num[round-mode=figures,round-precision=3]{164.21} & \num[round-mode=figures,round-precision=3]{107.547} $\pm$ \num[round-mode=figures,round-precision=3]{63.7469} & \num[round-mode=figures,round-precision=3]{8.52575} $\pm$ \num[round-mode=figures,round-precision=3]{8.0648} \\ 
\hline 
  Total  & \num[round-mode=figures,round-precision=3]{121060} $\pm$ \num[round-mode=figures,round-precision=3]{2157.23} & \num[round-mode=figures,round-precision=3]{11294.8} $\pm$ \num[round-mode=figures,round-precision=3]{284.411} & \num[round-mode=figures,round-precision=3]{180.903} $\pm$ \num[round-mode=figures,round-precision=3]{23.0855} & \num[round-mode=figures,round-precision=3]{58830.7} $\pm$ \num[round-mode=figures,round-precision=3]{1376.45} & \num[round-mode=figures,round-precision=3]{11689.4} $\pm$ \num[round-mode=figures,round-precision=3]{361.345} & \num[round-mode=figures,round-precision=3]{465.026} $\pm$ \num[round-mode=figures,round-precision=3]{29.0489} & \num[round-mode=figures,round-precision=3]{35997.5} $\pm$ \num[round-mode=figures,round-precision=3]{1273.84} & \num[round-mode=figures,round-precision=3]{14883.3} $\pm$ \num[round-mode=figures,round-precision=3]{502.308} & \num[round-mode=figures,round-precision=3]{1360.95} $\pm$ \num[round-mode=figures,round-precision=3]{71.384} \\ 
\hline 
  Data   & 120572 & 11275 & 176 & 58557 & 11707 & 466 & 35886 & 14877 & 1335 \\ 
\hline 
\end{tabular} 
\caption{Yields of the analysis} 
\end{center} 
\end{table} 
\end{document}