### 2018-06-30 figures/Hbb/fit/base_ebv2_signal_cH_default_unblinded_splusb/Tables/Yields_postFit.tex

\documentclass[10pt]{article}
\usepackage{siunitx}
\usepackage[margin=0.1in,landscape,papersize={210mm,350mm}]{geometry}
\begin{document}
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline 
 & 4jex-2bex & 4jex-3bex & 4jex-4bin & 5jex-2bex & 5jex-3bex & 5jex-4bin & 6jin-2bex & 6jin-3bex & 6jin-4bin\\
\hline 
  t#bar{t}+light   & \num[round-mode=figures,round-precision=3]{82845.9} $\pm$ \num[round-mode=figures,round-precision=3]{4191.85} & \num[round-mode=figures,round-precision=3]{4876.76} $\pm$ \num[round-mode=figures,round-precision=3]{501.636} & \num[round-mode=figures,round-precision=3]{15.5023} $\pm$ \num[round-mode=figures,round-precision=3]{12.3028} & \num[round-mode=figures,round-precision=3]{38012.6} $\pm$ \num[round-mode=figures,round-precision=3]{3128.71} & \num[round-mode=figures,round-precision=3]{3483.9} $\pm$ \num[round-mode=figures,round-precision=3]{466.698} & \num[round-mode=figures,round-precision=3]{15.8022} $\pm$ \num[round-mode=figures,round-precision=3]{9.49199} & \num[round-mode=figures,round-precision=3]{20096.7} $\pm$ \num[round-mode=figures,round-precision=3]{2531.65} & \num[round-mode=figures,round-precision=3]{2562.53} $\pm$ \num[round-mode=figures,round-precision=3]{491.271} & \num[round-mode=figures,round-precision=3]{21.4468} $\pm$ \num[round-mode=figures,round-precision=3]{23.4984} \\ 
  t#bar{t}+cc   & \num[round-mode=figures,round-precision=3]{11400} $\pm$ \num[round-mode=figures,round-precision=3]{4887.9} & \num[round-mode=figures,round-precision=3]{1360.1} $\pm$ \num[round-mode=figures,round-precision=3]{565.564} & \num[round-mode=figures,round-precision=3]{5.94047} $\pm$ \num[round-mode=figures,round-precision=3]{4.25326} & \num[round-mode=figures,round-precision=3]{8311.77} $\pm$ \num[round-mode=figures,round-precision=3]{3520.68} & \num[round-mode=figures,round-precision=3]{2000.89} $\pm$ \num[round-mode=figures,round-precision=3]{780.569} & \num[round-mode=figures,round-precision=3]{39.4621} $\pm$ \num[round-mode=figures,round-precision=3]{18.4738} & \num[round-mode=figures,round-precision=3]{7750.5} $\pm$ \num[round-mode=figures,round-precision=3]{3352.71} & \num[round-mode=figures,round-precision=3]{2974.4} $\pm$ \num[round-mode=figures,round-precision=3]{1173.97} & \num[round-mode=figures,round-precision=3]{58.8447} $\pm$ \num[round-mode=figures,round-precision=3]{25.671} \\ 
  t#bar{t}+bb   & \num[round-mode=figures,round-precision=3]{4284.48} $\pm$ \num[round-mode=figures,round-precision=3]{592.115} & \num[round-mode=figures,round-precision=3]{3403.26} $\pm$ \num[round-mode=figures,round-precision=3]{353.025} & \num[round-mode=figures,round-precision=3]{109.946} $\pm$ \num[round-mode=figures,round-precision=3]{16.8869} & \num[round-mode=figures,round-precision=3]{3415.95} $\pm$ \num[round-mode=figures,round-precision=3]{467.024} & \num[round-mode=figures,round-precision=3]{4917.61} $\pm$ \num[round-mode=figures,round-precision=3]{460.154} & \num[round-mode=figures,round-precision=3]{355.879} $\pm$ \num[round-mode=figures,round-precision=3]{28.7979} & \num[round-mode=figures,round-precision=3]{3396.38} $\pm$ \num[round-mode=figures,round-precision=3]{482.853} & \num[round-mode=figures,round-precision=3]{7517.33} $\pm$ \num[round-mode=figures,round-precision=3]{766.8} & \num[round-mode=figures,round-precision=3]{1105.78} $\pm$ \num[round-mode=figures,round-precision=3]{83.3295} \\ 
  topEW   & \num[round-mode=figures,round-precision=3]{173.513} $\pm$ \num[round-mode=figures,round-precision=3]{27.6992} & \num[round-mode=figures,round-precision=3]{35.0108} $\pm$ \num[round-mode=figures,round-precision=3]{5.8841} & \num[round-mode=figures,round-precision=3]{2.68504} $\pm$ \num[round-mode=figures,round-precision=3]{0.54471} & \num[round-mode=figures,round-precision=3]{167.542} $\pm$ \num[round-mode=figures,round-precision=3]{25.9028} & \num[round-mode=figures,round-precision=3]{65.4134} $\pm$ \num[round-mode=figures,round-precision=3]{10.4831} & \num[round-mode=figures,round-precision=3]{8.1585} $\pm$ \num[round-mode=figures,round-precision=3]{1.43202} & \num[round-mode=figures,round-precision=3]{212.274} $\pm$ \num[round-mode=figures,round-precision=3]{34.0437} & \num[round-mode=figures,round-precision=3]{146.736} $\pm$ \num[round-mode=figures,round-precision=3]{23.793} & \num[round-mode=figures,round-precision=3]{26.9001} $\pm$ \num[round-mode=figures,round-precision=3]{4.77355} \\ 
  ttH   & \num[round-mode=figures,round-precision=3]{62.6374} $\pm$ \num[round-mode=figures,round-precision=3]{7.83319} & \num[round-mode=figures,round-precision=3]{47.2558} $\pm$ \num[round-mode=figures,round-precision=3]{6.24357} & \num[round-mode=figures,round-precision=3]{4.67124} $\pm$ \num[round-mode=figures,round-precision=3]{0.684293} & \num[round-mode=figures,round-precision=3]{81.125} $\pm$ \num[round-mode=figures,round-precision=3]{8.92534} & \num[round-mode=figures,round-precision=3]{98.6353} $\pm$ \num[round-mode=figures,round-precision=3]{11.6012} & \num[round-mode=figures,round-precision=3]{16.5663} $\pm$ \num[round-mode=figures,round-precision=3]{2.29956} & \num[round-mode=figures,round-precision=3]{134.283} $\pm$ \num[round-mode=figures,round-precision=3]{15.464} & \num[round-mode=figures,round-precision=3]{239.948} $\pm$ \num[round-mode=figures,round-precision=3]{29.4611} & \num[round-mode=figures,round-precision=3]{61.4684} $\pm$ \num[round-mode=figures,round-precision=3]{8.75464} \\ 
  W+jetslight   & \num[round-mode=figures,round-precision=3]{6.1336} $\pm$ \num[round-mode=figures,round-precision=3]{6.31334} & \num[round-mode=figures,round-precision=3]{0.0116485} $\pm$ \num[round-mode=figures,round-precision=3]{0.0131938} & \num[round-mode=figures,round-precision=3]{4.05355e-06} $\pm$ \num[round-mode=figures,round-precision=3]{4.99067e-06} & \num[round-mode=figures,round-precision=3]{2.43781} $\pm$ \num[round-mode=figures,round-precision=3]{2.37932} & \num[round-mode=figures,round-precision=3]{0.0082768} $\pm$ \num[round-mode=figures,round-precision=3]{0.00937337} & \num[round-mode=figures,round-precision=3]{5.90626e-06} $\pm$ \num[round-mode=figures,round-precision=3]{7.22549e-06} & \num[round-mode=figures,round-precision=3]{1.14104} $\pm$ \num[round-mode=figures,round-precision=3]{1.07961} & \num[round-mode=figures,round-precision=3]{0.00748122} $\pm$ \num[round-mode=figures,round-precision=3]{0.0085961} & \num[round-mode=figures,round-precision=3]{9.93564e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.24269e-05} \\ 
  W+jetscharm   & \num[round-mode=figures,round-precision=3]{116.205} $\pm$ \num[round-mode=figures,round-precision=3]{59.336} & \num[round-mode=figures,round-precision=3]{1.23128} $\pm$ \num[round-mode=figures,round-precision=3]{0.746056} & \num[round-mode=figures,round-precision=3]{0.0024402} $\pm$ \num[round-mode=figures,round-precision=3]{0.00177021} & \num[round-mode=figures,round-precision=3]{47.3914} $\pm$ \num[round-mode=figures,round-precision=3]{23.9413} & \num[round-mode=figures,round-precision=3]{0.724028} $\pm$ \num[round-mode=figures,round-precision=3]{0.430586} & \num[round-mode=figures,round-precision=3]{0.00225932} $\pm$ \num[round-mode=figures,round-precision=3]{0.00162516} & \num[round-mode=figures,round-precision=3]{21.4785} $\pm$ \num[round-mode=figures,round-precision=3]{11.1851} & \num[round-mode=figures,round-precision=3]{0.613245} $\pm$ \num[round-mode=figures,round-precision=3]{0.367519} & \num[round-mode=figures,round-precision=3]{0.00343852} $\pm$ \num[round-mode=figures,round-precision=3]{0.0025106} \\ 
  W+jetsbeauty   & \num[round-mode=figures,round-precision=3]{4726.21} $\pm$ \num[round-mode=figures,round-precision=3]{1836.34} & \num[round-mode=figures,round-precision=3]{260.722} $\pm$ \num[round-mode=figures,round-precision=3]{104.695} & \num[round-mode=figures,round-precision=3]{2.88981} $\pm$ \num[round-mode=figures,round-precision=3]{1.34734} & \num[round-mode=figures,round-precision=3]{2047.29} $\pm$ \num[round-mode=figures,round-precision=3]{818.158} & \num[round-mode=figures,round-precision=3]{169.025} $\pm$ \num[round-mode=figures,round-precision=3]{68.0819} & \num[round-mode=figures,round-precision=3]{6.02499} $\pm$ \num[round-mode=figures,round-precision=3]{2.82343} & \num[round-mode=figures,round-precision=3]{1182.94} $\pm$ \num[round-mode=figures,round-precision=3]{472.32} & \num[round-mode=figures,round-precision=3]{183.637} $\pm$ \num[round-mode=figures,round-precision=3]{75.3516} & \num[round-mode=figures,round-precision=3]{12.5256} $\pm$ \num[round-mode=figures,round-precision=3]{5.727} \\ 
  Z+jetslight   & \num[round-mode=figures,round-precision=3]{1.6136} $\pm$ \num[round-mode=figures,round-precision=3]{1.55208} & \num[round-mode=figures,round-precision=3]{0.00375869} $\pm$ \num[round-mode=figures,round-precision=3]{0.00416768} & \num[round-mode=figures,round-precision=3]{1.39758e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.65619e-06} & \num[round-mode=figures,round-precision=3]{0.660099} $\pm$ \num[round-mode=figures,round-precision=3]{0.631076} & \num[round-mode=figures,round-precision=3]{0.00250452} $\pm$ \num[round-mode=figures,round-precision=3]{0.00281062} & \num[round-mode=figures,round-precision=3]{1.85667e-06} $\pm$ \num[round-mode=figures,round-precision=3]{2.27538e-06} & \num[round-mode=figures,round-precision=3]{0.301938} $\pm$ \num[round-mode=figures,round-precision=3]{0.272929} & \num[round-mode=figures,round-precision=3]{0.00216152} $\pm$ \num[round-mode=figures,round-precision=3]{0.00245911} & \num[round-mode=figures,round-precision=3]{2.97083e-06} $\pm$ \num[round-mode=figures,round-precision=3]{3.69504e-06} \\ 
  Z+jetscharm   & \num[round-mode=figures,round-precision=3]{20.4352} $\pm$ \num[round-mode=figures,round-precision=3]{10.3151} & \num[round-mode=figures,round-precision=3]{0.194919} $\pm$ \num[round-mode=figures,round-precision=3]{0.117146} & \num[round-mode=figures,round-precision=3]{0.000335712} $\pm$ \num[round-mode=figures,round-precision=3]{0.000245841} & \num[round-mode=figures,round-precision=3]{8.4355} $\pm$ \num[round-mode=figures,round-precision=3]{4.26127} & \num[round-mode=figures,round-precision=3]{0.128524} $\pm$ \num[round-mode=figures,round-precision=3]{0.0768793} & \num[round-mode=figures,round-precision=3]{0.000462529} $\pm$ \num[round-mode=figures,round-precision=3]{0.000332929} & \num[round-mode=figures,round-precision=3]{4.48254} $\pm$ \num[round-mode=figures,round-precision=3]{2.27427} & \num[round-mode=figures,round-precision=3]{0.126082} $\pm$ \num[round-mode=figures,round-precision=3]{0.0751513} & \num[round-mode=figures,round-precision=3]{0.000769429} $\pm$ \num[round-mode=figures,round-precision=3]{0.000545109} \\ 
  Z+jetsbeauty   & \num[round-mode=figures,round-precision=3]{1855.81} $\pm$ \num[round-mode=figures,round-precision=3]{726.854} & \num[round-mode=figures,round-precision=3]{102.616} $\pm$ \num[round-mode=figures,round-precision=3]{41.3097} & \num[round-mode=figures,round-precision=3]{1.93486} $\pm$ \num[round-mode=figures,round-precision=3]{1.03531} & \num[round-mode=figures,round-precision=3]{697.466} $\pm$ \num[round-mode=figures,round-precision=3]{274.516} & \num[round-mode=figures,round-precision=3]{74.4819} $\pm$ \num[round-mode=figures,round-precision=3]{30.3853} & \num[round-mode=figures,round-precision=3]{5.61523} $\pm$ \num[round-mode=figures,round-precision=3]{3.25619} & \num[round-mode=figures,round-precision=3]{352.092} $\pm$ \num[round-mode=figures,round-precision=3]{146.124} & \num[round-mode=figures,round-precision=3]{56.4242} $\pm$ \num[round-mode=figures,round-precision=3]{23.7334} & \num[round-mode=figures,round-precision=3]{3.4745} $\pm$ \num[round-mode=figures,round-precision=3]{2.17451} \\ 
  Single-top   & \num[round-mode=figures,round-precision=3]{6381.22} $\pm$ \num[round-mode=figures,round-precision=3]{885.113} & \num[round-mode=figures,round-precision=3]{399.607} $\pm$ \num[round-mode=figures,round-precision=3]{71.404} & \num[round-mode=figures,round-precision=3]{8.18061} $\pm$ \num[round-mode=figures,round-precision=3]{3.39484} & \num[round-mode=figures,round-precision=3]{2576.53} $\pm$ \num[round-mode=figures,round-precision=3]{548.887} & \num[round-mode=figures,round-precision=3]{325.298} $\pm$ \num[round-mode=figures,round-precision=3]{72.9207} & \num[round-mode=figures,round-precision=3]{13.5136} $\pm$ \num[round-mode=figures,round-precision=3]{4.20447} & \num[round-mode=figures,round-precision=3]{1239.51} $\pm$ \num[round-mode=figures,round-precision=3]{364.164} & \num[round-mode=figures,round-precision=3]{320.996} $\pm$ \num[round-mode=figures,round-precision=3]{98.2455} & \num[round-mode=figures,round-precision=3]{27.779} $\pm$ \num[round-mode=figures,round-precision=3]{10.1227} \\ 
  Dibosons   & \num[round-mode=figures,round-precision=3]{241.784} $\pm$ \num[round-mode=figures,round-precision=3]{84.0429} & \num[round-mode=figures,round-precision=3]{16.2441} $\pm$ \num[round-mode=figures,round-precision=3]{5.71612} & \num[round-mode=figures,round-precision=3]{0.495344} $\pm$ \num[round-mode=figures,round-precision=3]{0.222611} & \num[round-mode=figures,round-precision=3]{111.156} $\pm$ \num[round-mode=figures,round-precision=3]{47.8556} & \num[round-mode=figures,round-precision=3]{12.4395} $\pm$ \num[round-mode=figures,round-precision=3]{5.43722} & \num[round-mode=figures,round-precision=3]{0.762558} $\pm$ \num[round-mode=figures,round-precision=3]{0.391318} & \num[round-mode=figures,round-precision=3]{81.8407} $\pm$ \num[round-mode=figures,round-precision=3]{40.0188} & \num[round-mode=figures,round-precision=3]{16.6495} $\pm$ \num[round-mode=figures,round-precision=3]{8.19213} & \num[round-mode=figures,round-precision=3]{1.69053} $\pm$ \num[round-mode=figures,round-precision=3]{0.900296} \\ 
  SM-4tops   & \num[round-mode=figures,round-precision=3]{0.102591} $\pm$ \num[round-mode=figures,round-precision=3]{0.0384966} & \num[round-mode=figures,round-precision=3]{0.0969286} $\pm$ \num[round-mode=figures,round-precision=3]{0.0370062} & \num[round-mode=figures,round-precision=3]{0.0075197} $\pm$ \num[round-mode=figures,round-precision=3]{0.00617004} & \num[round-mode=figures,round-precision=3]{0.370247} $\pm$ \num[round-mode=figures,round-precision=3]{0.131963} & \num[round-mode=figures,round-precision=3]{0.492893} $\pm$ \num[round-mode=figures,round-precision=3]{0.158354} & \num[round-mode=figures,round-precision=3]{0.0541635} $\pm$ \num[round-mode=figures,round-precision=3]{0.0203327} & \num[round-mode=figures,round-precision=3]{5.14211} $\pm$ \num[round-mode=figures,round-precision=3]{1.54412} & \num[round-mode=figures,round-precision=3]{19.1277} $\pm$ \num[round-mode=figures,round-precision=3]{5.77236} & \num[round-mode=figures,round-precision=3]{4.95833} $\pm$ \num[round-mode=figures,round-precision=3]{1.54097} \\ 
  QCDE   & \num[round-mode=figures,round-precision=3]{7764.72} $\pm$ \num[round-mode=figures,round-precision=3]{3429.04} & \num[round-mode=figures,round-precision=3]{604.567} $\pm$ \num[round-mode=figures,round-precision=3]{233.932} & \num[round-mode=figures,round-precision=3]{20.0741} $\pm$ \num[round-mode=figures,round-precision=3]{14.1034} & \num[round-mode=figures,round-precision=3]{2645.66} $\pm$ \num[round-mode=figures,round-precision=3]{942.842} & \num[round-mode=figures,round-precision=3]{488.552} $\pm$ \num[round-mode=figures,round-precision=3]{225.005} & \num[round-mode=figures,round-precision=3]{3.60926} $\pm$ \num[round-mode=figures,round-precision=3]{4.81359} & \num[round-mode=figures,round-precision=3]{1224.61} $\pm$ \num[round-mode=figures,round-precision=3]{505.51} & \num[round-mode=figures,round-precision=3]{750.291} $\pm$ \num[round-mode=figures,round-precision=3]{341.548} & \num[round-mode=figures,round-precision=3]{28.4377} $\pm$ \num[round-mode=figures,round-precision=3]{24.6942} \\ 
  QCDMU   & \num[round-mode=figures,round-precision=3]{1249.61} $\pm$ \num[round-mode=figures,round-precision=3]{635.577} & \num[round-mode=figures,round-precision=3]{211.163} $\pm$ \num[round-mode=figures,round-precision=3]{100.134} & \num[round-mode=figures,round-precision=3]{9.03359} $\pm$ \num[round-mode=figures,round-precision=3]{8.47382} & \num[round-mode=figures,round-precision=3]{736.919} $\pm$ \num[round-mode=figures,round-precision=3]{369.678} & \num[round-mode=figures,round-precision=3]{71.0056} $\pm$ \num[round-mode=figures,round-precision=3]{44.6986} & \num[round-mode=figures,round-precision=3]{2e-05} $\pm$ \num[round-mode=figures,round-precision=3]{1.00366e-05} & \num[round-mode=figures,round-precision=3]{307.769} $\pm$ \num[round-mode=figures,round-precision=3]{164.21} & \num[round-mode=figures,round-precision=3]{107.601} $\pm$ \num[round-mode=figures,round-precision=3]{63.7808} & \num[round-mode=figures,round-precision=3]{8.52704} $\pm$ \num[round-mode=figures,round-precision=3]{8.06522} \\ 
  t#rightarrow Hc   & \num[round-mode=figures,round-precision=3]{-41.3493} $\pm$ \num[round-mode=figures,round-precision=3]{2.19773} & \num[round-mode=figures,round-precision=3]{-25.154} $\pm$ \num[round-mode=figures,round-precision=3]{1.71134} & \num[round-mode=figures,round-precision=3]{-0.519194} $\pm$ \num[round-mode=figures,round-precision=3]{0.149156} & \num[round-mode=figures,round-precision=3]{-25.4783} $\pm$ \num[round-mode=figures,round-precision=3]{2.09826} & \num[round-mode=figures,round-precision=3]{-18.9684} $\pm$ \num[round-mode=figures,round-precision=3]{1.89817} & \num[round-mode=figures,round-precision=3]{-0.555763} $\pm$ \num[round-mode=figures,round-precision=3]{0.153039} & \num[round-mode=figures,round-precision=3]{-15.0656} $\pm$ \num[round-mode=figures,round-precision=3]{1.66578} & \num[round-mode=figures,round-precision=3]{-12.4407} $\pm$ \num[round-mode=figures,round-precision=3]{1.72397} & \num[round-mode=figures,round-precision=3]{-1.17541} $\pm$ \num[round-mode=figures,round-precision=3]{0.234336} \\ 
\hline 
  Total  & \num[round-mode=figures,round-precision=3]{121130} $\pm$ \num[round-mode=figures,round-precision=3]{2239.1} & \num[round-mode=figures,round-precision=3]{11318.8} $\pm$ \num[round-mode=figures,round-precision=3]{413.449} & \num[round-mode=figures,round-precision=3]{181.363} $\pm$ \num[round-mode=figures,round-precision=3]{24.0046} & \num[round-mode=figures,round-precision=3]{58863.3} $\pm$ \num[round-mode=figures,round-precision=3]{1460.61} & \num[round-mode=figures,round-precision=3]{11708.6} $\pm$ \num[round-mode=figures,round-precision=3]{438.063} & \num[round-mode=figures,round-precision=3]{465.451} $\pm$ \num[round-mode=figures,round-precision=3]{29.8603} & \num[round-mode=figures,round-precision=3]{36011.4} $\pm$ \num[round-mode=figures,round-precision=3]{1318.36} & \num[round-mode=figures,round-precision=3]{14896.4} $\pm$ \num[round-mode=figures,round-precision=3]{550.482} & \num[round-mode=figures,round-precision=3]{1361.84} $\pm$ \num[round-mode=figures,round-precision=3]{76.7609} \\ 
\hline 
  Data   & 120572 & 11275 & 176 & 58557 & 11707 & 466 & 35886 & 14877 & 1335 \\ 
\hline 
\end{tabular} 
\caption{Yields of the analysis} 
\end{center} 
\end{table} 
\end{document}