### 2018-06-30 figures/Hbb/fit/base_ebv2_signal_uH_default_unblinded/Tables/Yields_postFit.tex

\documentclass[10pt]{article}
\usepackage{siunitx}
\usepackage[margin=0.1in,landscape,papersize={210mm,350mm}]{geometry}
\begin{document}
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline 
 & 4jex-2bex & 4jex-3bex & 4jex-4bin & 5jex-2bex & 5jex-3bex & 5jex-4bin & 6jin-2bex & 6jin-3bex & 6jin-4bin\\
\hline 
  t#bar{t}+light   & \num[round-mode=figures,round-precision=3]{82736.7} $\pm$ \num[round-mode=figures,round-precision=3]{3818.29} & \num[round-mode=figures,round-precision=3]{4870.48} $\pm$ \num[round-mode=figures,round-precision=3]{447.527} & \num[round-mode=figures,round-precision=3]{15.4586} $\pm$ \num[round-mode=figures,round-precision=3]{12.0169} & \num[round-mode=figures,round-precision=3]{37914.3} $\pm$ \num[round-mode=figures,round-precision=3]{2736.89} & \num[round-mode=figures,round-precision=3]{3470.01} $\pm$ \num[round-mode=figures,round-precision=3]{391.763} & \num[round-mode=figures,round-precision=3]{15.8632} $\pm$ \num[round-mode=figures,round-precision=3]{9.43766} & \num[round-mode=figures,round-precision=3]{20004.4} $\pm$ \num[round-mode=figures,round-precision=3]{2231.26} & \num[round-mode=figures,round-precision=3]{2549.13} $\pm$ \num[round-mode=figures,round-precision=3]{421.22} & \num[round-mode=figures,round-precision=3]{20.8411} $\pm$ \num[round-mode=figures,round-precision=3]{18.6847} \\ 
  t#bar{t}+cc   & \num[round-mode=figures,round-precision=3]{11455.7} $\pm$ \num[round-mode=figures,round-precision=3]{4842.46} & \num[round-mode=figures,round-precision=3]{1357.71} $\pm$ \num[round-mode=figures,round-precision=3]{568.799} & \num[round-mode=figures,round-precision=3]{5.85958} $\pm$ \num[round-mode=figures,round-precision=3]{4.13801} & \num[round-mode=figures,round-precision=3]{8375.44} $\pm$ \num[round-mode=figures,round-precision=3]{3441.7} & \num[round-mode=figures,round-precision=3]{2002.26} $\pm$ \num[round-mode=figures,round-precision=3]{785.096} & \num[round-mode=figures,round-precision=3]{39.3117} $\pm$ \num[round-mode=figures,round-precision=3]{18.3893} & \num[round-mode=figures,round-precision=3]{7831.2} $\pm$ \num[round-mode=figures,round-precision=3]{3251.3} & \num[round-mode=figures,round-precision=3]{2987.79} $\pm$ \num[round-mode=figures,round-precision=3]{1175.41} & \num[round-mode=figures,round-precision=3]{58.7913} $\pm$ \num[round-mode=figures,round-precision=3]{25.5942} \\ 
  t#bar{t}+bb   & \num[round-mode=figures,round-precision=3]{4258.72} $\pm$ \num[round-mode=figures,round-precision=3]{574.914} & \num[round-mode=figures,round-precision=3]{3395.21} $\pm$ \num[round-mode=figures,round-precision=3]{352.791} & \num[round-mode=figures,round-precision=3]{109.833} $\pm$ \num[round-mode=figures,round-precision=3]{16.9058} & \num[round-mode=figures,round-precision=3]{3399.68} $\pm$ \num[round-mode=figures,round-precision=3]{457.628} & \num[round-mode=figures,round-precision=3]{4910.58} $\pm$ \num[round-mode=figures,round-precision=3]{460.487} & \num[round-mode=figures,round-precision=3]{355.573} $\pm$ \num[round-mode=figures,round-precision=3]{28.7152} & \num[round-mode=figures,round-precision=3]{3383.78} $\pm$ \num[round-mode=figures,round-precision=3]{479.152} & \num[round-mode=figures,round-precision=3]{7504.21} $\pm$ \num[round-mode=figures,round-precision=3]{761.754} & \num[round-mode=figures,round-precision=3]{1104.67} $\pm$ \num[round-mode=figures,round-precision=3]{81.3061} \\ 
  topEW   & \num[round-mode=figures,round-precision=3]{173.264} $\pm$ \num[round-mode=figures,round-precision=3]{27.3871} & \num[round-mode=figures,round-precision=3]{34.9273} $\pm$ \num[round-mode=figures,round-precision=3]{5.71128} & \num[round-mode=figures,round-precision=3]{2.6856} $\pm$ \num[round-mode=figures,round-precision=3]{0.540473} & \num[round-mode=figures,round-precision=3]{167.597} $\pm$ \num[round-mode=figures,round-precision=3]{25.9071} & \num[round-mode=figures,round-precision=3]{65.3758} $\pm$ \num[round-mode=figures,round-precision=3]{10.4063} & \num[round-mode=figures,round-precision=3]{8.17168} $\pm$ \num[round-mode=figures,round-precision=3]{1.43466} & \num[round-mode=figures,round-precision=3]{212.973} $\pm$ \num[round-mode=figures,round-precision=3]{33.7948} & \num[round-mode=figures,round-precision=3]{147.078} $\pm$ \num[round-mode=figures,round-precision=3]{23.8292} & \num[round-mode=figures,round-precision=3]{27.0128} $\pm$ \num[round-mode=figures,round-precision=3]{4.76994} \\ 
  ttH   & \num[round-mode=figures,round-precision=3]{62.4596} $\pm$ \num[round-mode=figures,round-precision=3]{7.54464} & \num[round-mode=figures,round-precision=3]{47.1617} $\pm$ \num[round-mode=figures,round-precision=3]{6.06109} & \num[round-mode=figures,round-precision=3]{4.66843} $\pm$ \num[round-mode=figures,round-precision=3]{0.673201} & \num[round-mode=figures,round-precision=3]{81.0522} $\pm$ \num[round-mode=figures,round-precision=3]{8.86567} & \num[round-mode=figures,round-precision=3]{98.6344} $\pm$ \num[round-mode=figures,round-precision=3]{11.548} & \num[round-mode=figures,round-precision=3]{16.5886} $\pm$ \num[round-mode=figures,round-precision=3]{2.29912} & \num[round-mode=figures,round-precision=3]{134.656} $\pm$ \num[round-mode=figures,round-precision=3]{15.1917} & \num[round-mode=figures,round-precision=3]{240.832} $\pm$ \num[round-mode=figures,round-precision=3]{28.9758} & \num[round-mode=figures,round-precision=3]{61.7553} $\pm$ \num[round-mode=figures,round-precision=3]{8.68497} \\ 
  W+jetslight   & \num[round-mode=figures,round-precision=3]{6.13687} $\pm$ \num[round-mode=figures,round-precision=3]{6.3726} & \num[round-mode=figures,round-precision=3]{0.0116191} $\pm$ \num[round-mode=figures,round-precision=3]{0.0132552} & \num[round-mode=figures,round-precision=3]{4.02838e-06} $\pm$ \num[round-mode=figures,round-precision=3]{4.9843e-06} & \num[round-mode=figures,round-precision=3]{2.4412} $\pm$ \num[round-mode=figures,round-precision=3]{2.40463} & \num[round-mode=figures,round-precision=3]{0.0082668} $\pm$ \num[round-mode=figures,round-precision=3]{0.0094287} & \num[round-mode=figures,round-precision=3]{5.88125e-06} $\pm$ \num[round-mode=figures,round-precision=3]{7.23165e-06} & \num[round-mode=figures,round-precision=3]{1.14673} $\pm$ \num[round-mode=figures,round-precision=3]{1.09747} & \num[round-mode=figures,round-precision=3]{0.00749165} $\pm$ \num[round-mode=figures,round-precision=3]{0.00868194} & \num[round-mode=figures,round-precision=3]{9.92478e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.24923e-05} \\ 
  W+jetscharm   & \num[round-mode=figures,round-precision=3]{115.072} $\pm$ \num[round-mode=figures,round-precision=3]{56.4961} & \num[round-mode=figures,round-precision=3]{1.21345} $\pm$ \num[round-mode=figures,round-precision=3]{0.668815} & \num[round-mode=figures,round-precision=3]{0.00239716} $\pm$ \num[round-mode=figures,round-precision=3]{0.00149026} & \num[round-mode=figures,round-precision=3]{46.939} $\pm$ \num[round-mode=figures,round-precision=3]{22.8654} & \num[round-mode=figures,round-precision=3]{0.714879} $\pm$ \num[round-mode=figures,round-precision=3]{0.393686} & \num[round-mode=figures,round-precision=3]{0.00222412} $\pm$ \num[round-mode=figures,round-precision=3]{0.00142098} & \num[round-mode=figures,round-precision=3]{21.3729} $\pm$ \num[round-mode=figures,round-precision=3]{11.0021} & \num[round-mode=figures,round-precision=3]{0.607396} $\pm$ \num[round-mode=figures,round-precision=3]{0.344118} & \num[round-mode=figures,round-precision=3]{0.00339688} $\pm$ \num[round-mode=figures,round-precision=3]{0.00226103} \\ 
  W+jetsbeauty   & \num[round-mode=figures,round-precision=3]{4699.39} $\pm$ \num[round-mode=figures,round-precision=3]{1831.01} & \num[round-mode=figures,round-precision=3]{259.182} $\pm$ \num[round-mode=figures,round-precision=3]{104.289} & \num[round-mode=figures,round-precision=3]{2.87623} $\pm$ \num[round-mode=figures,round-precision=3]{1.34429} & \num[round-mode=figures,round-precision=3]{2039.18} $\pm$ \num[round-mode=figures,round-precision=3]{817.243} & \num[round-mode=figures,round-precision=3]{168.126} $\pm$ \num[round-mode=figures,round-precision=3]{67.9351} & \num[round-mode=figures,round-precision=3]{6.00435} $\pm$ \num[round-mode=figures,round-precision=3]{2.82184} & \num[round-mode=figures,round-precision=3]{1179.09} $\pm$ \num[round-mode=figures,round-precision=3]{471.719} & \num[round-mode=figures,round-precision=3]{182.957} $\pm$ \num[round-mode=figures,round-precision=3]{75.3672} & \num[round-mode=figures,round-precision=3]{12.5032} $\pm$ \num[round-mode=figures,round-precision=3]{5.73171} \\ 
  Z+jetslight   & \num[round-mode=figures,round-precision=3]{1.61319} $\pm$ \num[round-mode=figures,round-precision=3]{1.56458} & \num[round-mode=figures,round-precision=3]{0.00372087} $\pm$ \num[round-mode=figures,round-precision=3]{0.00409099} & \num[round-mode=figures,round-precision=3]{1.37802e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.61883e-06} & \num[round-mode=figures,round-precision=3]{0.660407} $\pm$ \num[round-mode=figures,round-precision=3]{0.636966} & \num[round-mode=figures,round-precision=3]{0.00248097} $\pm$ \num[round-mode=figures,round-precision=3]{0.00276278} & \num[round-mode=figures,round-precision=3]{1.83077e-06} $\pm$ \num[round-mode=figures,round-precision=3]{2.221e-06} & \num[round-mode=figures,round-precision=3]{0.302377} $\pm$ \num[round-mode=figures,round-precision=3]{0.275755} & \num[round-mode=figures,round-precision=3]{0.00214711} $\pm$ \num[round-mode=figures,round-precision=3]{0.0024307} & \num[round-mode=figures,round-precision=3]{2.94255e-06} $\pm$ \num[round-mode=figures,round-precision=3]{3.59597e-06} \\ 
  Z+jetscharm   & \num[round-mode=figures,round-precision=3]{20.2296} $\pm$ \num[round-mode=figures,round-precision=3]{9.78978} & \num[round-mode=figures,round-precision=3]{0.192183} $\pm$ \num[round-mode=figures,round-precision=3]{0.10577} & \num[round-mode=figures,round-precision=3]{0.000329513} $\pm$ \num[round-mode=figures,round-precision=3]{0.000209217} & \num[round-mode=figures,round-precision=3]{8.34363} $\pm$ \num[round-mode=figures,round-precision=3]{4.04684} & \num[round-mode=figures,round-precision=3]{0.12686} $\pm$ \num[round-mode=figures,round-precision=3]{0.0702107} & \num[round-mode=figures,round-precision=3]{0.000454394} $\pm$ \num[round-mode=figures,round-precision=3]{0.000283009} & \num[round-mode=figures,round-precision=3]{4.44878} $\pm$ \num[round-mode=figures,round-precision=3]{2.20319} & \num[round-mode=figures,round-precision=3]{0.124529} $\pm$ \num[round-mode=figures,round-precision=3]{0.0689165} & \num[round-mode=figures,round-precision=3]{0.000757345} $\pm$ \num[round-mode=figures,round-precision=3]{0.000476136} \\ 
  Z+jetsbeauty   & \num[round-mode=figures,round-precision=3]{1847.09} $\pm$ \num[round-mode=figures,round-precision=3]{725.378} & \num[round-mode=figures,round-precision=3]{102.151} $\pm$ \num[round-mode=figures,round-precision=3]{41.2591} & \num[round-mode=figures,round-precision=3]{1.92588} $\pm$ \num[round-mode=figures,round-precision=3]{1.03349} & \num[round-mode=figures,round-precision=3]{694.193} $\pm$ \num[round-mode=figures,round-precision=3]{274.143} & \num[round-mode=figures,round-precision=3]{73.9958} $\pm$ \num[round-mode=figures,round-precision=3]{30.2384} & \num[round-mode=figures,round-precision=3]{5.54591} $\pm$ \num[round-mode=figures,round-precision=3]{3.2085} & \num[round-mode=figures,round-precision=3]{351.493} $\pm$ \num[round-mode=figures,round-precision=3]{146.045} & \num[round-mode=figures,round-precision=3]{56.299} $\pm$ \num[round-mode=figures,round-precision=3]{23.7949} & \num[round-mode=figures,round-precision=3]{3.53004} $\pm$ \num[round-mode=figures,round-precision=3]{2.14947} \\ 
  Single-top   & \num[round-mode=figures,round-precision=3]{6405.21} $\pm$ \num[round-mode=figures,round-precision=3]{881.129} & \num[round-mode=figures,round-precision=3]{399.573} $\pm$ \num[round-mode=figures,round-precision=3]{70.1827} & \num[round-mode=figures,round-precision=3]{8.14191} $\pm$ \num[round-mode=figures,round-precision=3]{3.29931} & \num[round-mode=figures,round-precision=3]{2591.95} $\pm$ \num[round-mode=figures,round-precision=3]{545.578} & \num[round-mode=figures,round-precision=3]{326.041} $\pm$ \num[round-mode=figures,round-precision=3]{73.0191} & \num[round-mode=figures,round-precision=3]{13.575} $\pm$ \num[round-mode=figures,round-precision=3]{4.21531} & \num[round-mode=figures,round-precision=3]{1250.63} $\pm$ \num[round-mode=figures,round-precision=3]{360.943} & \num[round-mode=figures,round-precision=3]{322.922} $\pm$ \num[round-mode=figures,round-precision=3]{98.8107} & \num[round-mode=figures,round-precision=3]{28.0382} $\pm$ \num[round-mode=figures,round-precision=3]{10.1727} \\ 
  Dibosons   & \num[round-mode=figures,round-precision=3]{242.9} $\pm$ \num[round-mode=figures,round-precision=3]{84.1208} & \num[round-mode=figures,round-precision=3]{16.2734} $\pm$ \num[round-mode=figures,round-precision=3]{5.72528} & \num[round-mode=figures,round-precision=3]{0.495815} $\pm$ \num[round-mode=figures,round-precision=3]{0.222927} & \num[round-mode=figures,round-precision=3]{111.823} $\pm$ \num[round-mode=figures,round-precision=3]{47.877} & \num[round-mode=figures,round-precision=3]{12.4987} $\pm$ \num[round-mode=figures,round-precision=3]{5.45537} & \num[round-mode=figures,round-precision=3]{0.766919} $\pm$ \num[round-mode=figures,round-precision=3]{0.392579} & \num[round-mode=figures,round-precision=3]{82.3516} $\pm$ \num[round-mode=figures,round-precision=3]{40.0722} & \num[round-mode=figures,round-precision=3]{16.7367} $\pm$ \num[round-mode=figures,round-precision=3]{8.21993} & \num[round-mode=figures,round-precision=3]{1.70822} $\pm$ \num[round-mode=figures,round-precision=3]{0.904067} \\ 
  QCDE   & \num[round-mode=figures,round-precision=3]{7782.29} $\pm$ \num[round-mode=figures,round-precision=3]{3435.12} & \num[round-mode=figures,round-precision=3]{600.09} $\pm$ \num[round-mode=figures,round-precision=3]{229.152} & \num[round-mode=figures,round-precision=3]{20.0874} $\pm$ \num[round-mode=figures,round-precision=3]{14.1079} & \num[round-mode=figures,round-precision=3]{2656.51} $\pm$ \num[round-mode=figures,round-precision=3]{937.488} & \num[round-mode=figures,round-precision=3]{488.718} $\pm$ \num[round-mode=figures,round-precision=3]{223.903} & \num[round-mode=figures,round-precision=3]{3.60632} $\pm$ \num[round-mode=figures,round-precision=3]{4.81286} & \num[round-mode=figures,round-precision=3]{1225.53} $\pm$ \num[round-mode=figures,round-precision=3]{505.444} & \num[round-mode=figures,round-precision=3]{749.333} $\pm$ \num[round-mode=figures,round-precision=3]{338.803} & \num[round-mode=figures,round-precision=3]{28.4986} $\pm$ \num[round-mode=figures,round-precision=3]{24.7199} \\ 
  QCDMU   & \num[round-mode=figures,round-precision=3]{1250.34} $\pm$ \num[round-mode=figures,round-precision=3]{636.1} & \num[round-mode=figures,round-precision=3]{210.002} $\pm$ \num[round-mode=figures,round-precision=3]{98.9055} & \num[round-mode=figures,round-precision=3]{9.032} $\pm$ \num[round-mode=figures,round-precision=3]{8.47603} & \num[round-mode=figures,round-precision=3]{737.524} $\pm$ \num[round-mode=figures,round-precision=3]{369.77} & \num[round-mode=figures,round-precision=3]{71.0643} $\pm$ \num[round-mode=figures,round-precision=3]{44.7127} & \num[round-mode=figures,round-precision=3]{2e-05} $\pm$ \num[round-mode=figures,round-precision=3]{1.00366e-05} & \num[round-mode=figures,round-precision=3]{307.805} $\pm$ \num[round-mode=figures,round-precision=3]{164.154} & \num[round-mode=figures,round-precision=3]{107.192} $\pm$ \num[round-mode=figures,round-precision=3]{63.6139} & \num[round-mode=figures,round-precision=3]{8.52658} $\pm$ \num[round-mode=figures,round-precision=3]{8.06508} \\ 
  SM-4tops   & \num[round-mode=figures,round-precision=3]{0.101944} $\pm$ \num[round-mode=figures,round-precision=3]{0.0378504} & \num[round-mode=figures,round-precision=3]{0.0964923} $\pm$ \num[round-mode=figures,round-precision=3]{0.0364342} & \num[round-mode=figures,round-precision=3]{0.00741156} $\pm$ \num[round-mode=figures,round-precision=3]{0.00592196} & \num[round-mode=figures,round-precision=3]{0.367117} $\pm$ \num[round-mode=figures,round-precision=3]{0.126677} & \num[round-mode=figures,round-precision=3]{0.491372} $\pm$ \num[round-mode=figures,round-precision=3]{0.156655} & \num[round-mode=figures,round-precision=3]{0.0541694} $\pm$ \num[round-mode=figures,round-precision=3]{0.0203208} & \num[round-mode=figures,round-precision=3]{5.14482} $\pm$ \num[round-mode=figures,round-precision=3]{1.54471} & \num[round-mode=figures,round-precision=3]{19.1582} $\pm$ \num[round-mode=figures,round-precision=3]{5.78108} & \num[round-mode=figures,round-precision=3]{4.97106} $\pm$ \num[round-mode=figures,round-precision=3]{1.54508} \\ 
\hline 
  Total  & \num[round-mode=figures,round-precision=3]{121057} $\pm$ \num[round-mode=figures,round-precision=3]{2195.97} & \num[round-mode=figures,round-precision=3]{11294.3} $\pm$ \num[round-mode=figures,round-precision=3]{287.414} & \num[round-mode=figures,round-precision=3]{181.074} $\pm$ \num[round-mode=figures,round-precision=3]{23.3765} & \num[round-mode=figures,round-precision=3]{58828} $\pm$ \num[round-mode=figures,round-precision=3]{1406.91} & \num[round-mode=figures,round-precision=3]{11688.6} $\pm$ \num[round-mode=figures,round-precision=3]{368.518} & \num[round-mode=figures,round-precision=3]{465.064} $\pm$ \num[round-mode=figures,round-precision=3]{29.1391} & \num[round-mode=figures,round-precision=3]{35996.3} $\pm$ \num[round-mode=figures,round-precision=3]{1309.57} & \num[round-mode=figures,round-precision=3]{14884.4} $\pm$ \num[round-mode=figures,round-precision=3]{511.15} & \num[round-mode=figures,round-precision=3]{1360.85} $\pm$ \num[round-mode=figures,round-precision=3]{70.7275} \\ 
\hline 
  Data   & 120572 & 11275 & 176 & 58557 & 11707 & 466 & 35886 & 14877 & 1335 \\ 
\hline 
\end{tabular} 
\caption{Yields of the analysis} 
\end{center} 
\end{table} 
\end{document}