### 2018-06-30 figures/Hbb/fit/base_ebv2_signal_uH_default_unblinded_splusb/Tables/Yields_postFit.tex

\documentclass[10pt]{article}
\usepackage{siunitx}
\usepackage[margin=0.1in,landscape,papersize={210mm,350mm}]{geometry}
\begin{document}
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline 
 & 4jex-2bex & 4jex-3bex & 4jex-4bin & 5jex-2bex & 5jex-3bex & 5jex-4bin & 6jin-2bex & 6jin-3bex & 6jin-4bin\\
\hline 
  t#bar{t}+light   & \num[round-mode=figures,round-precision=3]{82621.5} $\pm$ \num[round-mode=figures,round-precision=3]{4264.13} & \num[round-mode=figures,round-precision=3]{4858.7} $\pm$ \num[round-mode=figures,round-precision=3]{518.823} & \num[round-mode=figures,round-precision=3]{15.3465} $\pm$ \num[round-mode=figures,round-precision=3]{12.2475} & \num[round-mode=figures,round-precision=3]{37818.5} $\pm$ \num[round-mode=figures,round-precision=3]{3238.56} & \num[round-mode=figures,round-precision=3]{3454.78} $\pm$ \num[round-mode=figures,round-precision=3]{490.283} & \num[round-mode=figures,round-precision=3]{15.7777} $\pm$ \num[round-mode=figures,round-precision=3]{9.67883} & \num[round-mode=figures,round-precision=3]{19936.5} $\pm$ \num[round-mode=figures,round-precision=3]{2589.67} & \num[round-mode=figures,round-precision=3]{2534.76} $\pm$ \num[round-mode=figures,round-precision=3]{503.364} & \num[round-mode=figures,round-precision=3]{20.2684} $\pm$ \num[round-mode=figures,round-precision=3]{22.5954} \\ 
  t#bar{t}+cc   & \num[round-mode=figures,round-precision=3]{11506.9} $\pm$ \num[round-mode=figures,round-precision=3]{4904.11} & \num[round-mode=figures,round-precision=3]{1357.86} $\pm$ \num[round-mode=figures,round-precision=3]{559.086} & \num[round-mode=figures,round-precision=3]{5.84705} $\pm$ \num[round-mode=figures,round-precision=3]{4.13245} & \num[round-mode=figures,round-precision=3]{8423.38} $\pm$ \num[round-mode=figures,round-precision=3]{3544.87} & \num[round-mode=figures,round-precision=3]{2004.02} $\pm$ \num[round-mode=figures,round-precision=3]{769.903} & \num[round-mode=figures,round-precision=3]{39.2415} $\pm$ \num[round-mode=figures,round-precision=3]{18.243} & \num[round-mode=figures,round-precision=3]{7882.2} $\pm$ \num[round-mode=figures,round-precision=3]{3382.01} & \num[round-mode=figures,round-precision=3]{2994.49} $\pm$ \num[round-mode=figures,round-precision=3]{1159.14} & \num[round-mode=figures,round-precision=3]{58.6695} $\pm$ \num[round-mode=figures,round-precision=3]{25.2641} \\ 
  t#bar{t}+bb   & \num[round-mode=figures,round-precision=3]{4251.6} $\pm$ \num[round-mode=figures,round-precision=3]{582.791} & \num[round-mode=figures,round-precision=3]{3394.36} $\pm$ \num[round-mode=figures,round-precision=3]{349.278} & \num[round-mode=figures,round-precision=3]{109.819} $\pm$ \num[round-mode=figures,round-precision=3]{16.9307} & \num[round-mode=figures,round-precision=3]{3395.16} $\pm$ \num[round-mode=figures,round-precision=3]{461.812} & \num[round-mode=figures,round-precision=3]{4910.7} $\pm$ \num[round-mode=figures,round-precision=3]{454.328} & \num[round-mode=figures,round-precision=3]{355.411} $\pm$ \num[round-mode=figures,round-precision=3]{28.7838} & \num[round-mode=figures,round-precision=3]{3382.28} $\pm$ \num[round-mode=figures,round-precision=3]{477.663} & \num[round-mode=figures,round-precision=3]{7503.07} $\pm$ \num[round-mode=figures,round-precision=3]{755.024} & \num[round-mode=figures,round-precision=3]{1104.51} $\pm$ \num[round-mode=figures,round-precision=3]{81.7063} \\ 
  topEW   & \num[round-mode=figures,round-precision=3]{173.079} $\pm$ \num[round-mode=figures,round-precision=3]{27.5058} & \num[round-mode=figures,round-precision=3]{34.8528} $\pm$ \num[round-mode=figures,round-precision=3]{5.83154} & \num[round-mode=figures,round-precision=3]{2.68137} $\pm$ \num[round-mode=figures,round-precision=3]{0.544856} & \num[round-mode=figures,round-precision=3]{167.575} $\pm$ \num[round-mode=figures,round-precision=3]{25.9057} & \num[round-mode=figures,round-precision=3]{65.3101} $\pm$ \num[round-mode=figures,round-precision=3]{10.4542} & \num[round-mode=figures,round-precision=3]{8.17101} $\pm$ \num[round-mode=figures,round-precision=3]{1.4353} & \num[round-mode=figures,round-precision=3]{213.259} $\pm$ \num[round-mode=figures,round-precision=3]{34.1966} & \num[round-mode=figures,round-precision=3]{147.122} $\pm$ \num[round-mode=figures,round-precision=3]{23.8674} & \num[round-mode=figures,round-precision=3]{27.0436} $\pm$ \num[round-mode=figures,round-precision=3]{4.81305} \\ 
  ttH   & \num[round-mode=figures,round-precision=3]{62.3644} $\pm$ \num[round-mode=figures,round-precision=3]{7.68687} & \num[round-mode=figures,round-precision=3]{47.0985} $\pm$ \num[round-mode=figures,round-precision=3]{6.12803} & \num[round-mode=figures,round-precision=3]{4.66345} $\pm$ \num[round-mode=figures,round-precision=3]{0.676345} & \num[round-mode=figures,round-precision=3]{81.0067} $\pm$ \num[round-mode=figures,round-precision=3]{8.88862} & \num[round-mode=figures,round-precision=3]{98.5882} $\pm$ \num[round-mode=figures,round-precision=3]{11.5599} & \num[round-mode=figures,round-precision=3]{16.5847} $\pm$ \num[round-mode=figures,round-precision=3]{2.30085} & \num[round-mode=figures,round-precision=3]{134.824} $\pm$ \num[round-mode=figures,round-precision=3]{15.4613} & \num[round-mode=figures,round-precision=3]{241.158} $\pm$ \num[round-mode=figures,round-precision=3]{29.5737} & \num[round-mode=figures,round-precision=3]{61.8439} $\pm$ \num[round-mode=figures,round-precision=3]{8.85361} \\ 
  W+jetslight   & \num[round-mode=figures,round-precision=3]{6.13507} $\pm$ \num[round-mode=figures,round-precision=3]{6.3893} & \num[round-mode=figures,round-precision=3]{0.0116201} $\pm$ \num[round-mode=figures,round-precision=3]{0.0132907} & \num[round-mode=figures,round-precision=3]{4.02845e-06} $\pm$ \num[round-mode=figures,round-precision=3]{4.99603e-06} & \num[round-mode=figures,round-precision=3]{2.44131} $\pm$ \num[round-mode=figures,round-precision=3]{2.40995} & \num[round-mode=figures,round-precision=3]{0.00826541} $\pm$ \num[round-mode=figures,round-precision=3]{0.00945306} & \num[round-mode=figures,round-precision=3]{5.87917e-06} $\pm$ \num[round-mode=figures,round-precision=3]{7.24772e-06} & \num[round-mode=figures,round-precision=3]{1.14915} $\pm$ \num[round-mode=figures,round-precision=3]{1.10065} & \num[round-mode=figures,round-precision=3]{0.00749934} $\pm$ \num[round-mode=figures,round-precision=3]{0.00870701} & \num[round-mode=figures,round-precision=3]{9.93406e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.25256e-05} \\ 
  W+jetscharm   & \num[round-mode=figures,round-precision=3]{114.335} $\pm$ \num[round-mode=figures,round-precision=3]{58.9381} & \num[round-mode=figures,round-precision=3]{1.2012} $\pm$ \num[round-mode=figures,round-precision=3]{0.739442} & \num[round-mode=figures,round-precision=3]{0.00236534} $\pm$ \num[round-mode=figures,round-precision=3]{0.00177045} & \num[round-mode=figures,round-precision=3]{46.6404} $\pm$ \num[round-mode=figures,round-precision=3]{23.8845} & \num[round-mode=figures,round-precision=3]{0.708481} $\pm$ \num[round-mode=figures,round-precision=3]{0.4294} & \num[round-mode=figures,round-precision=3]{0.00219848} $\pm$ \num[round-mode=figures,round-precision=3]{0.00161836} & \num[round-mode=figures,round-precision=3]{21.3067} $\pm$ \num[round-mode=figures,round-precision=3]{11.1806} & \num[round-mode=figures,round-precision=3]{0.602915} $\pm$ \num[round-mode=figures,round-precision=3]{0.367103} & \num[round-mode=figures,round-precision=3]{0.0033632} $\pm$ \num[round-mode=figures,round-precision=3]{0.00250144} \\ 
  W+jetsbeauty   & \num[round-mode=figures,round-precision=3]{4703.24} $\pm$ \num[round-mode=figures,round-precision=3]{1820.68} & \num[round-mode=figures,round-precision=3]{259.059} $\pm$ \num[round-mode=figures,round-precision=3]{103.615} & \num[round-mode=figures,round-precision=3]{2.87629} $\pm$ \num[round-mode=figures,round-precision=3]{1.33878} & \num[round-mode=figures,round-precision=3]{2041.93} $\pm$ \num[round-mode=figures,round-precision=3]{813.152} & \num[round-mode=figures,round-precision=3]{168.089} $\pm$ \num[round-mode=figures,round-precision=3]{67.4743} & \num[round-mode=figures,round-precision=3]{6.00793} $\pm$ \num[round-mode=figures,round-precision=3]{2.81029} & \num[round-mode=figures,round-precision=3]{1181.48} $\pm$ \num[round-mode=figures,round-precision=3]{470.367} & \num[round-mode=figures,round-precision=3]{183.135} $\pm$ \num[round-mode=figures,round-precision=3]{74.9288} & \num[round-mode=figures,round-precision=3]{12.5225} $\pm$ \num[round-mode=figures,round-precision=3]{5.71541} \\ 
  Z+jetslight   & \num[round-mode=figures,round-precision=3]{1.61224} $\pm$ \num[round-mode=figures,round-precision=3]{1.56826} & \num[round-mode=figures,round-precision=3]{0.00371508} $\pm$ \num[round-mode=figures,round-precision=3]{0.00410231} & \num[round-mode=figures,round-precision=3]{1.37542e-06} $\pm$ \num[round-mode=figures,round-precision=3]{1.62263e-06} & \num[round-mode=figures,round-precision=3]{0.660309} $\pm$ \num[round-mode=figures,round-precision=3]{0.638369} & \num[round-mode=figures,round-precision=3]{0.00247802} $\pm$ \num[round-mode=figures,round-precision=3]{0.00277031} & \num[round-mode=figures,round-precision=3]{1.82807e-06} $\pm$ \num[round-mode=figures,round-precision=3]{2.22597e-06} & \num[round-mode=figures,round-precision=3]{0.302496} $\pm$ \num[round-mode=figures,round-precision=3]{0.276306} & \num[round-mode=figures,round-precision=3]{0.00214612} $\pm$ \num[round-mode=figures,round-precision=3]{0.002437} & \num[round-mode=figures,round-precision=3]{2.94066e-06} $\pm$ \num[round-mode=figures,round-precision=3]{3.60504e-06} \\ 
  Z+jetscharm   & \num[round-mode=figures,round-precision=3]{20.0931} $\pm$ \num[round-mode=figures,round-precision=3]{10.2875} & \num[round-mode=figures,round-precision=3]{0.190314} $\pm$ \num[round-mode=figures,round-precision=3]{0.116286} & \num[round-mode=figures,round-precision=3]{0.000325423} $\pm$ \num[round-mode=figures,round-precision=3]{0.000242886} & \num[round-mode=figures,round-precision=3]{8.2915} $\pm$ \num[round-mode=figures,round-precision=3]{4.22279} & \num[round-mode=figures,round-precision=3]{0.125729} $\pm$ \num[round-mode=figures,round-precision=3]{0.0764736} & \num[round-mode=figures,round-precision=3]{0.000448786} $\pm$ \num[round-mode=figures,round-precision=3]{0.000331408} & \num[round-mode=figures,round-precision=3]{4.4267} $\pm$ \num[round-mode=figures,round-precision=3]{2.2708} & \num[round-mode=figures,round-precision=3]{0.123439} $\pm$ \num[round-mode=figures,round-precision=3]{0.0747982} & \num[round-mode=figures,round-precision=3]{0.000748721} $\pm$ \num[round-mode=figures,round-precision=3]{0.000541698} \\ 
  Z+jetsbeauty   & \num[round-mode=figures,round-precision=3]{1849.44} $\pm$ \num[round-mode=figures,round-precision=3]{721.993} & \num[round-mode=figures,round-precision=3]{102.175} $\pm$ \num[round-mode=figures,round-precision=3]{40.9932} & \num[round-mode=figures,round-precision=3]{1.92736} $\pm$ \num[round-mode=figures,round-precision=3]{1.03118} & \num[round-mode=figures,round-precision=3]{694.989} $\pm$ \num[round-mode=figures,round-precision=3]{272.727} & \num[round-mode=figures,round-precision=3]{74.0053} $\pm$ \num[round-mode=figures,round-precision=3]{30.064} & \num[round-mode=figures,round-precision=3]{5.52952} $\pm$ \num[round-mode=figures,round-precision=3]{3.21096} & \num[round-mode=figures,round-precision=3]{352.457} $\pm$ \num[round-mode=figures,round-precision=3]{145.98} & \num[round-mode=figures,round-precision=3]{56.3764} $\pm$ \num[round-mode=figures,round-precision=3]{23.6708} & \num[round-mode=figures,round-precision=3]{3.56573} $\pm$ \num[round-mode=figures,round-precision=3]{2.21084} \\ 
  Single-top   & \num[round-mode=figures,round-precision=3]{6413.64} $\pm$ \num[round-mode=figures,round-precision=3]{895.418} & \num[round-mode=figures,round-precision=3]{398.794} $\pm$ \num[round-mode=figures,round-precision=3]{71.5816} & \num[round-mode=figures,round-precision=3]{8.09999} $\pm$ \num[round-mode=figures,round-precision=3]{3.38477} & \num[round-mode=figures,round-precision=3]{2597.72} $\pm$ \num[round-mode=figures,round-precision=3]{555.813} & \num[round-mode=figures,round-precision=3]{325.886} $\pm$ \num[round-mode=figures,round-precision=3]{73.1366} & \num[round-mode=figures,round-precision=3]{13.5907} $\pm$ \num[round-mode=figures,round-precision=3]{4.22921} & \num[round-mode=figures,round-precision=3]{1255.17} $\pm$ \num[round-mode=figures,round-precision=3]{370.047} & \num[round-mode=figures,round-precision=3]{323.317} $\pm$ \num[round-mode=figures,round-precision=3]{99.1294} & \num[round-mode=figures,round-precision=3]{28.0955} $\pm$ \num[round-mode=figures,round-precision=3]{10.2489} \\ 
  Dibosons   & \num[round-mode=figures,round-precision=3]{243.339} $\pm$ \num[round-mode=figures,round-precision=3]{84.6212} & \num[round-mode=figures,round-precision=3]{16.2643} $\pm$ \num[round-mode=figures,round-precision=3]{5.72764} & \num[round-mode=figures,round-precision=3]{0.495776} $\pm$ \num[round-mode=figures,round-precision=3]{0.223015} & \num[round-mode=figures,round-precision=3]{112.109} $\pm$ \num[round-mode=figures,round-precision=3]{48.238} & \num[round-mode=figures,round-precision=3]{12.5143} $\pm$ \num[round-mode=figures,round-precision=3]{5.46963} & \num[round-mode=figures,round-precision=3]{0.767999} $\pm$ \num[round-mode=figures,round-precision=3]{0.393524} & \num[round-mode=figures,round-precision=3]{82.5787} $\pm$ \num[round-mode=figures,round-precision=3]{40.3695} & \num[round-mode=figures,round-precision=3]{16.7647} $\pm$ \num[round-mode=figures,round-precision=3]{8.24942} & \num[round-mode=figures,round-precision=3]{1.71383} $\pm$ \num[round-mode=figures,round-precision=3]{0.911589} \\ 
  QCDE   & \num[round-mode=figures,round-precision=3]{7788.63} $\pm$ \num[round-mode=figures,round-precision=3]{3447.93} & \num[round-mode=figures,round-precision=3]{598.103} $\pm$ \num[round-mode=figures,round-precision=3]{229.471} & \num[round-mode=figures,round-precision=3]{20.0615} $\pm$ \num[round-mode=figures,round-precision=3]{14.1004} & \num[round-mode=figures,round-precision=3]{2670.37} $\pm$ \num[round-mode=figures,round-precision=3]{964.573} & \num[round-mode=figures,round-precision=3]{487.793} $\pm$ \num[round-mode=figures,round-precision=3]{224.323} & \num[round-mode=figures,round-precision=3]{3.60627} $\pm$ \num[round-mode=figures,round-precision=3]{4.81286} & \num[round-mode=figures,round-precision=3]{1223.88} $\pm$ \num[round-mode=figures,round-precision=3]{503.911} & \num[round-mode=figures,round-precision=3]{747.718} $\pm$ \num[round-mode=figures,round-precision=3]{339.266} & \num[round-mode=figures,round-precision=3]{28.5377} $\pm$ \num[round-mode=figures,round-precision=3]{24.7368} \\ 
  QCDMU   & \num[round-mode=figures,round-precision=3]{1250.71} $\pm$ \num[round-mode=figures,round-precision=3]{636.813} & \num[round-mode=figures,round-precision=3]{208.951} $\pm$ \num[round-mode=figures,round-precision=3]{98.668} & \num[round-mode=figures,round-precision=3]{9.02458} $\pm$ \num[round-mode=figures,round-precision=3]{8.47464} & \num[round-mode=figures,round-precision=3]{738.365} $\pm$ \num[round-mode=figures,round-precision=3]{370.942} & \num[round-mode=figures,round-precision=3]{71.1194} $\pm$ \num[round-mode=figures,round-precision=3]{44.7488} & \num[round-mode=figures,round-precision=3]{2e-05} $\pm$ \num[round-mode=figures,round-precision=3]{1.00366e-05} & \num[round-mode=figures,round-precision=3]{307.735} $\pm$ \num[round-mode=figures,round-precision=3]{164.175} & \num[round-mode=figures,round-precision=3]{107.198} $\pm$ \num[round-mode=figures,round-precision=3]{63.6214} & \num[round-mode=figures,round-precision=3]{8.52622} $\pm$ \num[round-mode=figures,round-precision=3]{8.06499} \\ 
  SM-4tops   & \num[round-mode=figures,round-precision=3]{0.101669} $\pm$ \num[round-mode=figures,round-precision=3]{0.0380598} & \num[round-mode=figures,round-precision=3]{0.0962671} $\pm$ \num[round-mode=figures,round-precision=3]{0.0365775} & \num[round-mode=figures,round-precision=3]{0.00732719} $\pm$ \num[round-mode=figures,round-precision=3]{0.00611273} & \num[round-mode=figures,round-precision=3]{0.365631} $\pm$ \num[round-mode=figures,round-precision=3]{0.128712} & \num[round-mode=figures,round-precision=3]{0.490356} $\pm$ \num[round-mode=figures,round-precision=3]{0.157228} & \num[round-mode=figures,round-precision=3]{0.054148} $\pm$ \num[round-mode=figures,round-precision=3]{0.0203205} & \num[round-mode=figures,round-precision=3]{5.14698} $\pm$ \num[round-mode=figures,round-precision=3]{1.54586} & \num[round-mode=figures,round-precision=3]{19.1663} $\pm$ \num[round-mode=figures,round-precision=3]{5.78682} & \num[round-mode=figures,round-precision=3]{4.97279} $\pm$ \num[round-mode=figures,round-precision=3]{1.54702} \\ 
  t#rightarrow Hu   & \num[round-mode=figures,round-precision=3]{31.8275} $\pm$ \num[round-mode=figures,round-precision=3]{1.79374} & \num[round-mode=figures,round-precision=3]{17.2168} $\pm$ \num[round-mode=figures,round-precision=3]{1.28268} & \num[round-mode=figures,round-precision=3]{0.338133} $\pm$ \num[round-mode=figures,round-precision=3]{0.235499} & \num[round-mode=figures,round-precision=3]{18.2536} $\pm$ \num[round-mode=figures,round-precision=3]{1.69042} & \num[round-mode=figures,round-precision=3]{13.8434} $\pm$ \num[round-mode=figures,round-precision=3]{1.26685} & \num[round-mode=figures,round-precision=3]{0.348373} $\pm$ \num[round-mode=figures,round-precision=3]{0.163693} & \num[round-mode=figures,round-precision=3]{10.2895} $\pm$ \num[round-mode=figures,round-precision=3]{1.50077} & \num[round-mode=figures,round-precision=3]{8.08807} $\pm$ \num[round-mode=figures,round-precision=3]{1.21445} & \num[round-mode=figures,round-precision=3]{0.513095} $\pm$ \num[round-mode=figures,round-precision=3]{0.150948} \\ 
\hline 
  Total  & \num[round-mode=figures,round-precision=3]{121007} $\pm$ \num[round-mode=figures,round-precision=3]{2261.29} & \num[round-mode=figures,round-precision=3]{11277.7} $\pm$ \num[round-mode=figures,round-precision=3]{431.065} & \num[round-mode=figures,round-precision=3]{180.853} $\pm$ \num[round-mode=figures,round-precision=3]{23.8174} & \num[round-mode=figures,round-precision=3]{58799.5} $\pm$ \num[round-mode=figures,round-precision=3]{1465.66} & \num[round-mode=figures,round-precision=3]{11674.1} $\pm$ \num[round-mode=figures,round-precision=3]{460.406} & \num[round-mode=figures,round-precision=3]{464.746} $\pm$ \num[round-mode=figures,round-precision=3]{30.2351} & \num[round-mode=figures,round-precision=3]{35984.7} $\pm$ \num[round-mode=figures,round-precision=3]{1312.26} & \num[round-mode=figures,round-precision=3]{14875} $\pm$ \num[round-mode=figures,round-precision=3]{551.112} & \num[round-mode=figures,round-precision=3]{1360.27} $\pm$ \num[round-mode=figures,round-precision=3]{74.7837} \\ 
\hline 
  Data   & 120572 & 11275 & 176 & 58557 & 11707 & 466 & 35886 & 14877 & 1335 \\ 
\hline 
\end{tabular} 
\caption{Yields of the analysis} 
\end{center} 
\end{table} 
\end{document}