### 2018-08-03 ANA-TOPQ-2017-07-PAPER-metadata.tex
some fixes to text

%-------------------------------------------------------------------------------
% This file contains the title, author and abstract.
% It also contains all relevant document numbers used by the different cover pages.
%-------------------------------------------------------------------------------

% Title
%\AtlasTitle{Search for flavor-changing neutral current $t\rightarrow Hq$ ($q=u,c$) decays in pp collisions at $\sqrt{s}$=13 TeV with the ATLAS detector }
\AtlasTitle{Search for top quark decays $t \to Hq$ with 36 fb$^{-1}$ of $pp$ collision data at $\sqrt{s}=13~\tev$ with the ATLAS detector}

% Draft version:
% Should be 1.0 for the first circulation, and 2.0 for the second circulation.
% If given, adds draft version on front page, a 'DRAFT' box on top of each other page, 
% and line numbers.
% Comment or remove in final version.
\AtlasVersion{0.4}
% Abstract - % directly after { is important for correct indentation
\AtlasAbstract{%
A search for flavour-changing neutral current decays of a top quark into an up-type quark ($q=u, c$) and the 
Standard Model Higgs boson, $t\to Hq$, is presented. 
The search is based on a dataset of $pp$ collisions at $\sqrt{s}=13~\tev$ recorded in 2015 and 2016 with the ATLAS detector at the 
CERN Large Hadron Collider and corresponds to an integrated luminosity of 36.1 fb$^{-1}$.
Two analyses are performed searching for top quark pair events in which one top quark decays to $Wb$ and the other top quark decays to $Hq$,
and targeting the $H \to b\bar{b}$ and $H \to \tau^+\tau^-$  decay modes, respectively.  
They exploit either the high multiplicity of $b$-quark jets, or the presence of hadronically decaying tau leptons, and employ different
multivariate techniques to separate the signal from the background, which is dominated by $t\bar{t}$ production.
No significant excess of events above the background expectation is found, and 95\% CL upper limits on the $t\to Hq$ branching ratios are derived.
The combination of these searches with ATLAS searches in diphoton and multilepton final states significantly improves the sensitivity, 
yielding observed (expected) 95\% CL upper limits on the $t\to Hc$ and $t\to Hu$ branching ratios of $1.1 \times 10^{-3}$ ($8.3 \times 10^{-4}$) 
and $1.2 \times 10^{-3}$ ($8.3 \times 10^{-4}$) respectively.
The corresponding combined observed (expected) upper limits on the $|\lambda_{tcH}|$ and $|\lambda_{tuH}|$ couplings are 0.064 (0.055) and 0.064 (0.055) respectively. 
%Upper limits in the $t\to Hc$ versus $t\to Hu$ branching ratio plane, as well as best-fit branching ratios, are also reported.
These are the most restrictive direct bounds on $tqH$ interactions measured so far.
}

% Author - this does not work with revtex (add it after \begin{document})
\author{The ATLAS Collaboration}

% ATLAS reference code, to help ATLAS members to locate the paper
\AtlasRefCode{TOPQ-2017-07}

% ATLAS date - arXiv submission; usually filled in by the Physics Office
% \AtlasDate{\today}

% ATLAS heading - heading at top of title page. Set for TDR etc.
%\AtlasHeading{ATLAS ABC TDR}

% Submission journal and final reference
\AtlasJournal{JHEP}
% \AtlasJournal{Phys.\ Lett.\ B.}
% \AtlasJournalRef{\PLB 789 (2017) 123}
% \AtlasDOI{}

 %-------------------------------------------------------------------------------
% The following information is needed for the cover page. The commands are only defined
% if you use the coverpage option in atlasdoc or use the atlascover package
%-------------------------------------------------------------------------------

% List of supporting notes  (leave as null \AtlasCoverSupportingNote{} if you want to skip this option)
\AtlasCoverSupportingNote{Search for $\ttbar \to WbHq$, $H \to b\bar{b}$}{https://cds.cern.ch/record/2257631}
\AtlasCoverSupportingNote{Search for $\ttbar \to WbHq$, $H \to \tau^+\tau^-$}{https://cds.cern.ch/record/2273683}
\AtlasCoverSupportingNote{Combination of $\ttbar \to WbHq$ searches}{https://cds.cern.ch/record/2312520/}
%
% OR (the 2nd option is deprecated, especially for CONF and PUB notes)
%
% Supporting material TWiki page  (leave as null \AtlasCoverTwikiURL{} if you want to skip this option)
% \AtlasCoverTwikiURL{https://twiki.cern.ch/twiki/bin/view/Atlas/WebHome}

% Comment deadline
% \AtlasCoverCommentsDeadline{DD Month 2016}

% Analysis team members - contact editors should no longer be specified
% as there is a generic email list name for the editors
\AtlasCoverAnalysisTeam{
$H \to b\bar{b}$: Trisha Farooque, Davide Gerbaudo, Aurelio Juste, Nicola Orlando,  \\ Cheng Peng,  Laura Pereira, 
Yulia Rodina, Yanjun Tu,  Lo\"ic Val\'ery, \\ Tal Van Daalen, Jia-Shian Wang, Daiki Yamaguchi; \\
$H \to \tau^+\tau^-$: Xin Chen, Antonio De Maria, Boyang Li, Ligang Xia; \\
Combination: Peter Onyisi, Harish Potti}

% Editorial Board Members - indicate the Chair by a (chair) after his/her name
% Give either all members at once (then they appear on one line), or separately
\AtlasCoverEdBoardMember{Ketevi Assamagan~(chair), Garabed Halladjian, Fabrice Hubaut, Michele Weber~(chair)}

% \AtlasCoverEdBoardMember{EdBoard~Chair~(chair)}
% \AtlasCoverEdBoardMember{EB~Member~1}
% \AtlasCoverEdBoardMember{EB~Member~2}
% \AtlasCoverEdBoardMember{EB~Member~3}

% Editors egroup
\AtlasCoverEgroupEditors{atlas-TOPQ-2017-07-editors@cern.ch}

% EdBoard egroup
\AtlasCoverEgroupEdBoard{atlas-TOPQ-2017-07-editorial-board@cern.ch}