### 2018-10-03 ANA-TOPQ-2017-07-PAPER-auxmat.tex
porting language editor comments

\part*{Auxiliary material}
\addcontentsline{toc}{part}{Auxiliary material}
%-------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[htbp]
\begin{center}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex2bex_lep0_pt.eps}}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex2bex_met.eps}} \\
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex2bex_hthad.eps}} 
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex2bex_mbb_mindR.eps}} \\ 
\caption{\small{$\Hbb$ search: Comparison between the data and background prediction for several kinematic 
distributions in the (4j, 2b) region (prior to the application of the cut on the LH discriminant above 0.6) before  the fit to data (``Pre-Fit''). 
The distributions are shown for (a) lepton $\pt$, (b) $\met$, (c) scalar sum of the transverse momenta of 
the jets ($\hthad$), and (d) the invariant mass of the two $b$-tagged jets with lowest 
$\Delta R$ separation ($\mbb$).
The small contributions from $\ttbar V$, $\ttbar H$, single top, $W/Z$+jets, diboson, and multijet backgrounds are combined 
into a single background source referred to as ``Non-$\ttbar$''. 
The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
added to the background prediction.
The last bin in all figures contains the overflow.
The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction. 
The blue triangles indicate points that are outside the vertical range of the figure. 
The hashed area represents the total uncertainty of the background, excluding the normalisation uncertainty of the $\ttbin$ background, 
which is determined via a likelihood fit to data.}}
\label{fig:Hbb_extravars_4j2b}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[htbp]
%\begin{center}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex3bex_lep0_pt.eps}}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex3bex_met.eps}} \\
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex3bex_hthad.eps}} 
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep4jex3bex_mbb_mindR.eps}} \\ 
%\caption{\small{$\Hbb$ search: Comparison between the data and background prediction for several kinematic 
%distributions in the (4j, 3b) region before performing the fit to data (``Pre-Fit''). 
%The distributions are shown for (a) lepton $\pt$, (b) $\met$, (c) scalar sum of the transverse momenta of 
%the jets ($\hthad$), and (d) the invariant mass distribution of the two $b$-tagged jets with lowest 
%$\Delta R$ separation ($\mbb$).
%The small contributions from $\ttbar V$, $\ttbar H$, single top, $W/Z$+jets, diboson, and multijet backgrounds are combined 
%into a single background source referred to as ``Non-$\ttbar$''. 
%The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
%added on top of the background prediction.
%The last bin in all figures contains the overflow.
%The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction. 
%The blue triangles indicate points that are outside the vertical range of the figure. 
%The hashed area represents the total uncertainty on the background, excluding the normalisation uncertainty of the $\ttbin$ background, 
%which is determined via a likelihood fit to data.}}
%\label{fig:Hbb_extravars_4j3b}
%\end{center}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[htbp]
\begin{center}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep5jex3bex_lep0_pt.eps}}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep5jex3bex_met.eps}} \\
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep5jex3bex_hthad.eps}} 
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Hbb/other_variables/c1lep5jex3bex_mbb_mindR.eps}} \\ 
\caption{\small{$\Hbb$ search: Comparison between the data and background prediction for several kinematic 
distributions in the (5j, 3b) region before the fit to data (``Pre-Fit''). 
The distributions are shown for (a) lepton $\pt$, (b) $\met$, (c) scalar sum of the transverse momenta of 
the jets ($\hthad$), and (d) the invariant mass of the two $b$-tagged jets with lowest 
$\Delta R$ separation ($\mbb$).
The small contributions from $\ttbar V$, $\ttbar H$, single top, $W/Z$+jets, diboson, and multijet backgrounds are combined 
into a single background source referred to as ``Non-$\ttbar$''. 
The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
added to the background prediction.
The last bin in all figures contains the overflow.
The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction. 
The blue triangles indicate points that are outside the vertical range of the figure. 
The hashed area represents the total uncertainty of the background, excluding the normalisation uncertainty of the $\ttbin$ background, 
which is determined via a likelihood fit to data.}}
\label{fig:Hbb_extravars_5j3b}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[t]
\begin{center}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/x1_fit_lephad_3j_FR.eps}}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/x1_fit_lephad_4j_FR.eps}} \\
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/x2_fit_lephad_3j_FR.eps}}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/x2_fit_lephad_4j_FR.eps}} \\
\caption{$\Htautau$ search: Comparison between the data and background prediction for the distribution of some of the most
discriminating BDT input variables in the $\lephad$ channel before the fit to data (``Pre-Fit''). The distributions are shown for
$x_{1}^{\text{fit}}$ in (a) the ($\lephad$, 3j) region and (b) the ($\lephad$, $\geq$4j) region, and for
$x_{2}^{\text{fit}}$ in (c) the ($\lephad$, 3j)  region and (d) the ($\lephad$, $\geq$4j) region.
The contributions with real $\had$ candidates from $\ttbar$,  $\ttbar V$, $\ttbar H$, and single top backgrounds are combined into
a single background source referred to as ``Top (real $\had$)", whereas the small contributions from 
$Z\to \ell^+\ell^-$ ($\ell = e, \mu$) and diboson backgrounds are combined into ``Other''. 
The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
added to the background prediction.
%The first and the last bins in all figures contain the underflow and overflow respectively.
The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction.
The hashed area represents the total uncertainty of the background, excluding the normalisation uncertainty of the fake $\had$ background, 
which is determined via a likelihood fit to data.}
\label{fig:BDT_inputs_lephad_2}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[t]
\begin{center}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/x2_fit_hadhad_3j_FR.eps}}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/x2_fit_hadhad_4j_FR.eps}} \\
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/ptL2_hadhad_3j_FR.eps}}
\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/Htautau/control_plots/ptL2_hadhad_4j_FR.eps}} \\
\caption{$\Htautau$ search: Comparison between the data and background prediction for the distribution of two of the most 
discriminating BDT input variables in the $\hadhad$ channel before the fit to data (``Pre-Fit''). The distributions are shown for
$x_{2}^{\text{fit}}$ in (a) the ($\hadhad$, 3j) region and (b) the ($\hadhad$, $\geq$4j) region, and for
$p_{\text{T},2}$ in (c) the ($\hadhad$, 3j)  region and (d) the ($\hadhad$, $\geq$4j) region.
The contributions with real $\had$ candidates from $\ttbar$,  $\ttbar V$, $\ttbar H$, and single top backgrounds are combined into
a single background source referred to as ``Top (real $\had$)", whereas the small contributions from 
$Z\to \ell^+\ell^-$ ($\ell = e, \mu$) and diboson backgrounds are combined into ``Other''. 
The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
added to the background prediction.
The first and the last bins in the figures in (c) and (d) contain the underflow and overflow respectively.
The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction.
The hashed area represents the total uncertainty of the background, excluding the normalisation uncertainty of the fake $\had$ background, 
which is determined via a likelihood fit to data.} 
\label{fig:BDT_inputs_hadhad_2}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[t]
%\begin{center}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}} \\
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}} \\
%\caption{$\Htautau$ search: Comparison between the data and background prediction for the distribution of two of the most 
%important BDT input variables in the $\hadhad$ channel before performing the fit to data (``Pre-Fit''). The distributions are shown for
%$x_{2}^{\text{fit}}$ in (a) the ($\hadhad$, 3j) region and (b) the ($\hadhad$, $\geq$4j) region, and for
%$p_{\text{T},2}$ in (c) the ($\hadhad$, 3j)  region and (d) the ($\hadhad$, $\geq$4j) region.
%The contributions with real $\had$ candidates from $\ttbar$,  $\ttbar V$, $\ttbar H$, and single top backgrounds are combined into
%a single background source referred to as ``Top (real $\had$)", whereas the small contributions from 
%$Z\to \ell^+\ell^-$ ($\ell = e, \mu$) and diboson backgrounds are combined into ``Other''. 
%The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
%added on top of the background prediction.
%The first and the last bins in the figures in (c) and (d) contain the underflow and overflow respectively.
%The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction.
%The hashed area represents the total uncertainty on the background, excluding the normalisation uncertainty of the fake $\had$ background, 
%which is determined via a likelihood fit to data.} 
%\label{fig:BDT_inputs_hadhad_3}
%\end{center}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[t]
%\begin{center}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}} \\
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}}
%\subfloat[]{\includegraphics[width=0.40\textwidth]{figures/placeholder8x8.eps}} \\
%\caption{$\Htautau$ search: Comparison between the data and background prediction for the distribution of two of the most 
%important BDT input variables in the $\hadhad$ channel before performing the fit to data (``Pre-Fit''). The distributions are shown for
%$x_{2}^{\text{fit}}$ in (a) the ($\hadhad$, 3j) region and (b) the ($\hadhad$, $\geq$4j) region, and for
%$p_{\text{T},2}$ in (c) the ($\hadhad$, 3j)  region and (d) the ($\hadhad$, $\geq$4j) region.
%The contributions with real $\had$ candidates from $\ttbar$,  $\ttbar V$, $\ttbar H$, and single top backgrounds are combined into
%a single background source referred to as ``Top (real $\had$)", whereas the small contributions from 
%$Z\to \ell^+\ell^-$ ($\ell = e, \mu$) and diboson backgrounds are combined into ``Other''. 
%The expected $\Hc$ signal (solid red) corresponding to $\BR(t\to Hc)=1\%$ is also shown,
%added on top of the background prediction.
%The first and the last bins in the figures in (c) and (d) contain the underflow and overflow respectively.
%The bottom panel displays the ratio of data to the SM background (``Bkg'') prediction.
%The hashed area represents the total uncertainty on the background, excluding the normalisation uncertainty of the fake $\had$ background, 
%which is determined via a likelihood fit to data.} 
%\label{fig:BDT_inputs_hadhad_3}
%\end{center}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%