### 2022-06-02 sections/conclusion.tex

%-------------------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
%-------------------------------------------------------------------------------
A search for flavour-changing neutral current processes involving a top quark, another up-type quark ($q=u, c$), and a SM Higgs boson is presented. FCNC $tqH$ interaction is sought both in the $t\bar{t}$ decay mode, where one top quark decays via SM processes and the other one decays through $t\rightarrow qH$, and in the production mode ($pp\rightarrow tH$) with a single top quark produced via the
FCNC interaction where $t\to Wb$.
%The search is based on a dataset of $pp$ collisions at $\sqrt{s}=13~\tev$ recorded with the ATLAS detector at the CERN Large Hadron Collider and corresponding to an integrated luminosity of 139 fb$^{-1}$. The search selects events with two hadronically decaying $\tau$-lepton candidates ($\tauhad$) or at least one $\tauhad$ with an additional lepton (e,$\mu$), a $b$-jet as well as multiple jets. Multivariate techniques based on BDT are used to distinguish the signal from the background on the basis of their different kinematics.
%The search sensitivities have been improved significantly compared to the previous analysis using 36$^{-1}$ of run-2 data~\cite{Aaboud2019SearchFT}. No significant excess of events above the background expectation is found.
A slight excess of data is observed above background with a significance of 2.3$\sigma$. Upper limits at the
95\% confidence level are set on the $t\to qH$ branching ratios and the corresponding
dimension-6 operator Wilson coefficients in the effective $tqH$ couplings. 
%about a factor of 2 improvement to 95\% CL upper limits on the $t\to Hq$ branching ratios are reported in Table~\ref{tab:limits_summary}, along with the Wilson coefficients with respect to the previous partial run-2 results~\cite{Aaboud2019SearchFT}.
Observed (expected) 95\% CL upper limits are set on the $t\to cH$ and $t\to uH$ branching ratios of $9.4\times10^{-4}$ ($4.8^{+2.2}_{-1.4}\times10^{-4}$) and $6.9\times10^{-4}$ ($3.5^{+1.5}_{-1.0}\times10^{-4}$).
%The observed (expected) upper limits on the D6 Wilson coefficients from the combination of the searches are set to $C_{c\phi}<1.38\,(0.97)$ and $C_{u\phi}<1.18\,(0.83)$, respectively.
The corresponding combined observed (expected) upper limits on the dimension-6 operator Wilson coefficients in
the effective $tqH$ couplings for a new physics scale $\Lambda$ of 1~TeV are $C_{c\phi} <1.35\, (0.97)$ and $C_{u\phi} <1.16\, (0.82)$, respectively.
%These results improve significantly upon previous ATLAS and CMS limits on the $t\rightarrow Hq$ branching ratios.
These results improve significantly upon the previously published ATLAS studies in this channel and provide more stringent limits than the ATLAS  $t\rightarrow qH$ combination~\cite{fcnc36}.