### 2022-06-02 sections/tau_FF_appendix.tex

\section{Tau fake scale factor calibration in the CRtt}
\label{sec:tauFF_appendix}

Figure~\ref{fig:ttCR_mtt} presents the post-fit distributions of the leading $\thad$ $\pt$ in the CRtt. These show good agreement between the data and the fitted background model.
Tables~\ref{tab:ff1_summary} and~\ref{tab:ff2_summary} summarize the fake scale factors of 1-prong and 3-prong taus
derived from the CRtt including both statistical and systematic uncertainties.

\begin{figure}[H]
\centering
\begin{tabular}{@{}ccc@{}}
\includegraphics[page=1,width=0.29\textwidth]{figures/ttCR/tuH_reg1l1tau2b2j_os_log_ttCR.pdf} &
\includegraphics[page=1,width=0.29\textwidth]{figures/ttCR/tuH_reg1l1tau2b2j_ss_log_ttCR.pdf}&
\includegraphics[page=1,width=0.29\textwidth]{figures/ttCR/tuH_reg1l1tau2b3j_os_log_ttCR.pdf}\\
(a) & (b) & (c) \\
\includegraphics[page=1,width=0.29\textwidth]{figures/ttCR/tuH_reg1l1tau2b3j_ss_log_ttCR.pdf}&
\includegraphics[page=1,width=0.29\textwidth]{figures/ttCR/tuH_reg2l1tau1bnj_log_ttCR.pdf}&
\includegraphics[page=1,width=0.29\textwidth]{figures/ttCR/tuH_reg2l1tau2bnj_log_ttCR.pdf}\\
(d) & (e) & (f)\\
\end{tabular}
\caption{Leading $\thad$ $\pt$  distributions obtained after the fit to data (``Post-Fit'') for the $\thad$ fake scale factors in the following CRtt:
  $t_{\ell}t_{h}$2b$\thad$-2j OS (a), $t_{\ell}t_{h}$2b$\thad$-2j SS (b), $t_{\ell}t_{h}$2b$\thad$-3j OS (c), $t_{\ell}t_{h}$2b$\thad$-3j SS (d),
  $t_{\ell}t_{\ell}$1b$\thad$-nj (e) and $t_{\ell}t_{\ell}$2b$\thad$-nj (f).
  The total size of the statistical and systematic uncertainties of the background prediction is indicated by the hatched band.
  Overflow events are included in the last bin. The lower panels show the data-to-prediction ratio.}
\label{fig:ttCR_mtt}
\end{figure}

\begin{table}
\caption{ Summary of fake tau (1-prong) scale factors derived in the CRtt. The numbers are shown as: central values, statistical uncertainties, and systematics uncertainties. }
\begin{center}
\begin{tabular}{lcccccc}
\toprule\toprule

Fake tau types & \multicolumn{3}{c}{1-prong tau}  \\ \midrule
$\thad$ $\pt$                                   &  $25-35$ GeV  & $35-45$ GeV  &  $45-$ GeV            \\
\midrule
Type-1                          &$0.71 \pm 0.01 \pm 0.03 $ &$0.61 \pm 0.02 \pm 0.04 $ &$0.38 \pm 0.02 \pm 0.05 $  \\
% W fake OS
Type-2                          &$0.76 \pm 0.06 \pm 0.04 $ & $0.37 \pm 0.08 \pm 0.02$ & $0.74 \pm 0.08 \pm 0.02 $ \\
% W fake SS
Type-3                          &$0.62 \pm 0.10 \pm 0.03 $ &$0.83 \pm 0.09 \pm 0.03 $ &$0.94 \pm 0.07 \pm 0.02 $  \\
%  b fake
Type-4                          &$1.20 \pm 0.02 \pm 0.01 $ & $1.01 \pm 0.04 \pm 0.02 $ &$0.76 \pm 0.03 \pm 0.03 $  \\
% other fake
\bottomrule\bottomrule
\end{tabular}
\label{tab:ff1_summary}
\end{center}
\end{table}

\begin{table}
\caption{ Summary of fake tau (3-prong) scale factors derived in the CRtt. The numbers are shown as: central values, statistical uncertainties, and systematics uncertainties. }
\begin{center}
\begin{tabular}{lcccccc}
\toprule\toprule

Fake tau types        & \multicolumn{3}{c}{3-prong tau}  \\ \midrule
$\thad$ $\pt$                             &  $25-35$ GeV  & $35-45$ GeV       &  $45-$ GeV  \\
\midrule
Type-1                                    & $1.01 \pm 0.03 \pm 0.04 $ & $1.09 \pm 0.04 \pm 0.05 $ & $0.30 \pm 0.05 \pm 0.07 $ \\
% W fake OS
Type-2                                    & $0.93 \pm 0.10 \pm 0.04 $ & $1.05 \pm 0.09 \pm 0.03 $ & $0.79 \pm 0.09 \pm 0.04 $ \\
% W fake SS
Type-3                                    & $1.07 \pm 0.13 \pm 0.03 $ &$1.39 \pm 0.12 \pm 0.03 $ &$1.26 \pm 0.10 \pm 0.04 $  \\
%  b fake
Type-4                                    &$1.28 \pm 0.07 \pm 0.02 $ &$0.66 \pm 0.08 \pm 0.01 $ & $0.71 \pm 0.07 \pm 0.02 $ \\
% other fake
\bottomrule\bottomrule
\end{tabular}
\label{tab:ff2_summary}
\end{center}
\end{table}