### 2022-06-29 sections/systematic_uncertainties.tex

%-------------------------------------------------------------------------------
\section{Systematic uncertainties}
\label{sec:systematics}
%-------------------------------------------------------------------------------

Several sources of systematic uncertainty that can affect the normalisation of signal 
and background and/or the shape of their corresponding discriminant distributions are considered.
Each source is considered to be uncorrelated with the other sources.  
Correlations of a given systematic uncertainty are maintained across processes and channels 
as appropriate. The following sections describe the systematic uncertainties considered.
Table~\ref{tab:had_sys_impact} shows a summary of the dominant systematic uncertainties in the measured $\BR(t\to qH)$ branching ratio
resulting from the fits to the data in the signal regions as described in Section~\ref{sec:result}.

\subsection{Luminosity}
\label{sec:syst_lumi}

The uncertainty in the integrated luminosity is 1.7\%, affecting the overall normalisation of all processes estimated from the simulation. 
It is derived, following a methodology similar to that detailed in Ref.~\cite{Aaboud:2016hhf} and using the LUCID-2 detector 
for the baseline luminosity measurements \cite{Avoni:2018iuv}, from a calibration of the luminosity scale using $x$--$y$ beam-separation scans.

\subsection{Reconstructed objects}
\label{sec:syst_objects}

Uncertainties associated with electrons, muons, and $\had$ candidates arise from the trigger, reconstruction,  
identification and, in the case of electrons and muons, isolation efficiencies, as well as the momentum scale and resolution. 
These are measured using $Z\to \ell^+\ell^-$ and $J/\psi\to \ell^+\ell^-$ events ($\ell =e, \mu$)~\cite{ATLAS-CONF-2016-024,Aad:2016jkr} 
in the case of electrons and muons, and using $Z\to \tau^+\tau^-$ events in the case of $\had$ candidates~\cite{ATLAS-CONF-2017-029}.

Uncertainties associated with jets arise from the jet energy scale (JES)
and resolution (JER), and the efficiency to pass the JVT requirements. 
The largest contribution comes from the jet energy scale, whose uncertainty dependence on jet $\pt$ and $\eta$, jet flavour, and pile-up treatment
is split into 43 uncorrelated components that are treated independently~\cite{Aaboud:2017jcu}. The total JES uncertainty is
below 5\% for most jets and below 1\% for central jets with $\pt$ between 300~\GeV\ and 2~\TeV. The difference between the JER values
in data and MC events is represented by one nuisance parameter (NP). It is applied to the MC events by smearing the jet $\pt$ within the prescribed uncertainty.

Uncertainties associated with energy scales and resolutions of leptons and jets 
are propagated to $\met$. Additional uncertainties originating from the modelling 
of the underlying event, in particular its impact on the $\pt$ scale and resolution 
of unclustered energy, are negligible.

Efficiencies to tag $b$-jets and $c$-jets in the simulation are corrected by $\pt$-dependent factors to match the efficiencies in data,
while the light-jet efficiency is scaled by $\pt$- and $\eta$-dependent factors.
The $b$-jet efficiency is measured in a data sample enriched in $\ttbar$ events~\cite{Aad:2019epj79}, %EPJ C79(2019)970 {Aaboud:2018xwy},
  while the $c$-jet efficiency is measured
using $\ttbar$ events~\cite{ATLAS-CONF-2018-001} or $W$+$c$-jet events~\cite{Aad:2015ydr}. 
The light-jet efficiency is measured in a multi-jet data sample enriched in light-flavour jets~\cite{ATLAS-CONF-2018-006}.
The uncertainties in these scale factors come from a total of 44 independent sources affecting $b$-jets, 19 sources affecting $c$-jets, and 19 sources affecting light jets. 
These systematic uncertainties are taken to be uncorrelated between $b$-jets, $c$-jets, and light jets.
%The uncertainty on the pileup reweighing is evaluated by varying the pileup scale factors by 1$\sigma$ based on the reweighing of the average interactions per bunch crossing. However, this uncertainty is highly correlated with the luminosity uncertainty and may be an overestimate.
Scale factors are applied to reweight simulated events in order to obtain the pile-up distribution corresponding to the data.
An uncertainty in these reweighting scale factors is estimated by reweighting the profile in data while keeping it within its uncertainties.
%considered based on the disagreement between the instantaneous luminosity in data and in simulation.

\subsection{Background modelling}
\label{sec:syst_bkgmodeling}

A number of sources of systematic uncertainty affecting the modelling of $t\bar{t}$+jets are considered: the choice of renormalisation and factorisation scale in the matrix-element calculation, the choice of matching scale when matching the matrix elements to the parton shower generator, the uncertainty in the value of $\alpha_\text{s}$ when modelling initial-state radiation (ISR), and
the choice of renormalisation scale when modelling final-state radiation (FSR).
%%An uncertainty of  6\% is assigned to the inclusive $\ttbar$ production
%%cross section~\cite{Czakon:2011xx}, including contributions from varying the factorisation and renormalisation 
%%scales, as well as from the top-quark mass, the PDF and $\alpha_{\textrm{S}}$. The latter two represent the largest contribution 
%%to the overall theoretical uncertainty in the cross section and were calculated using the PDF4LHC prescription~\cite{Botje:2011sn} 
%%with the NNPDF3.0(NLO), CT10 NNLO~\cite{Lai:2010vv,Gao:2013xoa} and NNPDF2.3(LO) 5F FFN~\cite{Ball:2012cx} PDF sets.

The $h_{\mathrm{damp}}$ parameter, which controls the amount of radiation produced by the parton shower in
\powheg\,v2, is set to 1.5$m_t$ in the $t\bar t$ sample. An alternative sample was generated with $h_{\mathrm{damp}}=3m_t$. The
difference between the two samples is treated as a systematic uncertainty called `$t\bar{t}$ $h_{\mathrm{damp}}$'. 
%The uncertainty associated with the choice of NLO generator is derived by comparing the nominal prediction from
%{\powheg}+{\pythiaeight} with a prediction from \textsc{Sherpa}~2.2.1. For the latter, the matrix-element calculation is performed 
%for up to two partons at NLO and up to four partons at LO using \textsc{Comix} and \textsc{OpenLoops}, and
%merged with the {\sherpa} parton shower using the ME+PS@NLO prescription.

The uncertainty due to the choice of parton shower and hadronisation (PS \& Had) model is derived 
by comparing the predictions from {\powheg} interfaced either to \textsc{Pythia\,8} or \textsc{Herwig\,7}.
The latter uses the \textsc{MMHT2014lo}~\cite{Harland-Lang:2014zoa} PDF set in combination with the H7UE tune~\cite{Bellm:2015jjp}.
The uncertainty in the modelling of additional radiation from the PS is assessed by
varying the corresponding parameter of the A14 set~\cite{ATL-PHYS-PUB-2016-004} and by varying the radiation renormalisation and factorisation scales
by a factor of 2.0 and 0.5, respectively. 
%The uncertainty in the modelling of additional radiation is assessed with two alternative {\powheg}+{\pythiaeight} samples:
%a sample with increased radiation (referred to as radHi) is obtained by decreasing the renormalisation and factorisation scales  
%by a factor of two, %doubling the $h_{\textrm{damp}}$ parameter,
%and using the Var3c upward variation of the A14 parameter set;
%a sample with decreased radiation (referred to as radLow) is obtained by increasing the scales by a factor of two  %and using the Var3c downward variation of the A14 set~\cite{ATL-PHYS-PUB-2016-004}.

Another significant background in the hadronic channel stems from the $Z\rightarrow \tau\tau$ samples. Several sources of uncertainty are
considered for these samples: the PDF variation, which is evaluated by considering the standard deviation
produced by the event weights from the 100 NNPDF replicas for the \textsc{NNPDF3.0nnlo}~\cite{Ball:2015NNPDF} PDF set used in \sherpa,
the renormalisation ($\mu_\text{r}$) and factorisation ($\mu_\text{f}$) scales, the jet-to-parton matching uncertainty, the resummation scale uncertainty,
the variation in the choice of $\alpha_\text{s}$, and the use of alternative PDFs, which is evaluated by comparing predictions from the \textsc{NNPDF3.0nnlo} PDF set (nominal)
with those from the \textsc{MMHT2014nnlo68cl} and \textsc{CT14nnlo}~\cite{Lai:2010vv,Gao:2013xoa} PDF sets.

Uncertainties affecting the normalisation of the $V$+jets background are estimated separately for $V$+light-jets, $V$+$\geq$1$c$+jets,
and $V$+$\geq$1$b$+jets subprocesses. The total normalisation uncertainty of $V$+jets processes is estimated to be approximately 30\% by taking the maximum
difference between the data and the total background prediction in the different analysis regions considered, but requiring exactly zero, one, and two $b$-jets.
This is driven mainly by differences between MC simulations in the $V$+jets regions with a high multiplicity of jets.  
%Agreement between data and predicted background 
%in these modified regions, which are dominated by $V$+light-jets, is found to be within approximately 30\%. This bound is taken to 
%be the normalisation uncertainty, correlated across all $V$+jets subprocesses. 
%Since {\sherpa}~2.2 has been found to underestimate $V$+heavy-flavour production by about a factor
%of 1.3~\cite{Aaboud:2017xsd}, additional 30\% normalisation uncertainties are assumed for $V$+$\geq$1$c$+jets and $V$+$\geq$1$b$+jets
%subprocesses, considered uncorrelated between them.

%To estimate the uncertainty originating from ISR modelling in the single top samples, the same procedure as $t\bar{t}$ is used. One NP reflects the symmetrised effect of the two shower variations Var3cDown and Var3cUp, and the other NP with name ``scale'' describe the symmetrised effect from the independent variations of the renormalisation and factorisation scale. The impact of the uncertainty from FSR modelling is estimated by reweighing the nominal single top sample. The two variations $\mu^{FSR}_{R} \times 0.5$ and $\mu^{FSR}_{R} \times 2$ are considered.  The standard deviation of 100 NNPDF3.0nnlo variations is calculated Uncertainties on the PDF are evaluated following the recommended prescription in Ref.~\cite{ttRun2}.

%The values of the scale-parameters $\mu_{f}$ and $\mu_{f}$ are varied by factors of 2.0 and 0.5 with respect to their default values to obtain the uncertainties related to the renormalization and factorization scales in ttV samples. Uncertainties on the PDF are evaluated following the same procedure in Ref.~\cite{ttZRun2}.

%Following the recommendations of the PMG group, the uncertainties to be considered for diboson background include~\cite{dibosonRes} scale variations of the renormalization and factorization scale,
%PDF variations and $\alpha_{S}$ variations.

%A $+5.8\%-9.2\%$ normalization uncertainty is considered for the ttH background, corresponding to the scale and $\alpha_{S}$ uncertainties in the NLO cross-section computation. A constant PDF uncertainty of $\pm3.6$\% is assumed, as it was done for the previous measurement. Therefore, a overall systematics with variation 9\% is entered in the final fit model~\cite{ttZRun2}.

Uncertainties affecting the modelling of the single-top-quark background include a 
$+5\%$/$-4\%$ uncertainty in the total cross section, which is estimated as a weighted average 
of the theoretical uncertainties in $t$-, $tW$- and $s$-channel production~\cite{Kidonakis:2011wy,Kidonakis:2010ux,Kidonakis:2010tc}.
Additional uncertainties associated with the parton shower, hadronisation and ISR/FSR radiation are also considered by using the same procedure
as for $t\bar t$. Uncertainties in the diboson background normalisation include those estimated from variations of the renormalisation and factorisation scales,  \textsc{NNPDF3.0nnlo} variations and $\alpha_\text{s}$ variations. Uncertainties in the $\ttbar V$ and $\ttbar H$ cross sections are estimated to be 12\% and [$+5.8\%/-9.2\%$], respectively,
from the uncertainties in their respective NLO theoretical cross sections~\cite{ttZRun2}.

%by comparing the nominal
%%%samples with alternative samples where generator parameters are varied.
%%%For the $t$- and $tW$-channel processes, an uncertainty due to the choice of parton shower and hadronisation model is derived 
%%%by comparing events produced by {\powheg} interfaced to {\pythia}~8 or {\herwigpp}.
%%%These uncertainties are treated as fully correlated among single-top-quark production processes, but uncorrelated with the
%%%corresponding uncertainty of the $\ttbar$+jets background.
%%%An additional systematic uncertainty in $tW$-channel production concerning the separation 
%%%between $t\bar{t}$ and $tW$ at NLO is assessed by comparing
%%%the nominal sample, which uses the diagram removal scheme~\cite{Frixione:2008yi}, with an alternative sample
%%%using the diagram subtraction scheme~\cite{Frixione:2008yi}.

%%%%(it is assumed that two jets originate from the $W/Z$ decay, as in $WW/WZ \to \ell \nu jj$). 
%%%%Therefore, the total normalisation uncertainty is $5\% \oplus \sqrt{N-2}\times 24\%$, where $N$ is the selected jet multiplicity,  
%%%%resulting in 34\%, 42\%, and 48\%, for events with exactly 4 jets, exactly 5 jets, and $\geq$6 jets, respectively. 
%%%%Recent comparisons between data and {\sherpa}~2.1.1 for $WZ(\to \ell\nu\ell\ell) + \geq$4 jets show
%%%%agreement within the experimental uncertainty of approximately 40\%~\cite{Aaboud:2016yus}, which further justifies the above uncertainty.
%%%%Given the very small contribution of this background to the total prediction, the final result is not affected by the assumed modelling
%%%%uncertainties.
%%%%

The statistical uncertainties of the fake-$\tau$-background calibration in the leptonic channels are applied with uncorrelated uncertainties for
different sources of fake $\tau$-leptons and different $\pt$ slices. The uncertainties in the MC modelling of the various processes used for
evaluation of the fake factors in the CRtt are treated as fully correlated in the fit.
The uncertainty in the ABCD method is applied to the normalisation factors for muons and electrons, including both
its statistical fluctuation and the differences between the normalisation factors for the various leptonic channels.
The uncertainties in the fake-factor method applied to the hadronic channel includes the statistical uncertainty of each fake factor and
differences between the sets of fake factors derived from different signal-depleted CRs.
%The uncertainties on the MC modelling of the various processes used for
%evaluation of the fake factors in the CRtt's are treated as fully correlated in the fit.

\subsection{Signal modelling}
\label{sec:syst_sigmodeling}

Several normalisation and shape uncertainties are taken into account and are treated as fully correlated for the $\Hq$ and $pp\to tH$ signals.
Uncertainties in the Higgs boson branching ratios are taken
%from the measured values in Ref.~\cite{Zyla:2020zbs}.
into account by following the recommendation in Ref.~\cite{deFlorian:2016spz}.
The uncertainties due to ISR, FSR, the scales, and the PDFs, are considered and are treated as fully correlated in all SRs. The parton shower uncertainties are estimated by comparing
the nominal sample with an alternative sample interfaced with \textsc{Herwig\,7}.
%the predictions from {\powheg} interfaced either to {\pythiaeight} or {\herwig7}. 
%treated correlatedly with $\ttbar$ samples.

%\begin{table}[h!]
%\caption{List of relative uncertainties on the upper limits of $\BR(t\to qH)$ ($q=u,c$) obtained from the combined fit. The uncertainties are symmetrised for presentation and grouped into the categories described in the text.}
%\label{tab:had_sys_impact}
%\begin{center}
%\begin{tabular}{%
%@{}l%
%S%
%S%
%@{}%
%}
%\toprule\toprule
%\multirow{2}{*}{Source of uncertainty}      & \multicolumn{2}{c}{$\Delta\BR/\BR [\%]$} \\
%                                            & \multicolumn{1}{c}{$t\rightarrow uH$} & \multicolumn{1}{c}{$t\rightarrow cH$} \\\midrule
%Lepton                                  & 0.6           &1.0         \\
%Met                                     & 0.7           &0.8         \\
%Luminosity and Pileup                   & 1.0           &1.3         \\
%Theoretical uncertainty in other MC     & 2.1           &2.9         \\
%Jets                                    & 2.4           &3.2         \\
%Flavour tagging                         & 2.7           &3.7         \\
%Misidentified $\tau$                    & 3.4           &4.8         \\
%Theoretical uncertainty in $t\bar{t}$   & 2.9           &4.3         \\
%Hadronic $\tau$ decays                  & 3.3           &4.4         \\
%Theoretical uncertainty in signal       & 5.3           &7.0         \\\midrule
%Total statistics uncertainty            & 5.1           &7.0         \\
%Total systematic uncertainty            & 11.2          &15.4        \\ \midrule
%Total                                   & 12.3          &16.9        \\

%Lepton ID                               & 0.6           &1.0         \\
%$\met$                                  & 0.7           &0.8         \\
%Fake lepton  modeling                   & 0.9           &1.1         \\
%Luminosity and Pileup                   & 0.9           &1.3         \\
%Other MC modeling                       & 2.1           &2.9         \\
%JES and JER                             & 2.4           &3.2         \\
%Flavour tagging                         & 2.7           &3.7         \\
%$t\bar{t}$ modeling                     & 2.9           &4.3         \\
%Fake $\tau$ modeling                    & 3.2           &4.6         \\
%$\tau$ ID                               & 3.3           &4.4         \\
%Signal modeling including Br($H\to\tau\tau$)       & 5.3           &7.0         \\\midrule
%Total systematic uncertainty            & 11.2          &15.5        \\ 
%Total statistics uncertainty            & 5.1           &7.0         \\\midrule
%Total                                   & 12.3          &17.0        \\

%tuH:
%Fake LEP method    0.00919939  ( +0.00919939, -0.00919939 )
%Fake method    0.0323159  ( +0.0323159, -0.0323159 )
%FullSyst    0.112111  ( +0.112111, -0.112111 )
%Gammas    0.0509968  ( +0.0509968, -0.0509968 )
%Instrument    0.0093934  ( +0.0093934, -0.0093934 )
%JET    0.0235705  ( +0.0235705, -0.0235705 )
%Lepton    0.00568277  ( +0.00568277, -0.00568277 )
%MET    0.00704844  ( +0.00704844, -0.00704844 )
%SignalTheory    0.0526104  ( +0.0526104, -0.0526104 )
%TAU    0.032714  ( +0.032714, -0.032714 )
%Theory    0.0210686  ( +0.0210686, -0.0210686 )
%btag    0.027376  ( +0.027376, -0.027376 )
%ttbarTheory    0.0289194  ( +0.0289194, -0.0289194 )

%tcH:
%Fake LEP method    0.010914  ( +0.010914, -0.010914 )
%Fake method    0.0455017  ( +0.0455017, -0.0455017 )
%FullSyst    0.154686  ( +0.154686, -0.154686 )
%Gammas    0.0700708  ( +0.0700708, -0.0700708 )
%Instrument    0.0131647  ( +0.0131647, -0.0131647 )
%JET    0.0320897  ( +0.0320897, -0.0320897 )
%Lepton    0.0104552  ( +0.0104552, -0.0104552 )
%MET    0.00841447  ( +0.00841447, -0.00841447 )
%SignalTheory    0.0699833  ( +0.0699833, -0.0699833 )
%TAU    0.0437292  ( +0.0437292, -0.0437292 )
%Theory    0.029182  ( +0.029182, -0.029182 )
%btag    0.0367613  ( +0.0367613, -0.0367613 )
%ttbarTheory    0.0427391  ( +0.0427391, -0.0427391 )

%\bottomrule\bottomrule
%\end{tabular} 
%\end{center}
%\end{table}

%% Fake method    0.0476721  ( +0.0476721, -0.0476721 )
%% FullSyst    0.154686  ( +0.154686, -0.154686 )
%% Gammas    0.0700708  ( +0.0700708, -0.0700708 )
%% Instrument    0.0131647  ( +0.0131647, -0.0131647 )
%% JET    0.0320897  ( +0.0320897, -0.0320897 )
%% Lepton    0.0104552  ( +0.0104552, -0.0104552 )
%% MET    0.00841447  ( +0.00841447, -0.00841447 )
%% SignalTheory    0.0699833  ( +0.0699833, -0.0699833 )
%% TAU    0.0437292  ( +0.0437292, -0.0437292 )
%% Theory    0.029182  ( +0.029182, -0.029182 )
%% btag    0.0367613  ( +0.0367613, -0.0367613 )
%% ttbarTheory    0.0427391  ( +0.0427391, -0.0427391 )

%% Fake method   0.0341326  ( +0.0341326, -0.0341326 )
%% FullSyst      0.112111      ( +0.112111, -0.112111 )
%% Gammas        0.0509968       ( +0.0509968, -0.0509968 )
%% Instrument    0.0093934  ( +0.0093934, -0.0093934 )
%% JET           0.0235705  ( +0.0235705, -0.0235705 )
%% Lepton        0.00568277  ( +0.00568277, -0.00568277 )
%% MET           0.00704844  ( +0.00704844, -0.00704844 )
%% SignalTheory  0.0526104  ( +0.0526104, -0.0526104 )
%% TAU           0.032714  ( +0.032714, -0.032714 )
%% Theory        0.0210686  ( +0.0210686, -0.0210686 )
%% btag          0.027376  ( +0.027376, -0.027376 )
%% ttbarTheory   0.0289194  ( +0.0289194, -0.0289194 )

%\begin{table}[htbp]
%\caption{List of relative uncertainties of the signal strength from the combined fit in hadronic channel. The uncertainties are symmetrised for presentation and %grouped into the categories described in the text. %The quadrature sum of the individual uncertainties is not equal to the total uncertainty due to correlations %introduced by the fit
%}
%\small
%\centering
%\begin{tabular}{cc} \toprule\toprule
%$\text{Uncertainty}    $   & $\Delta\mu/\mu[\%]$ \\\midrule
%$\text{Fake modelling} $   & $13.5$ \\
%$\text{Instrument}     $   & $2.6$  \\
%$\text{Jets}           $   & $1.7$   \\
%$\text{Met}            $   & $0.4$   \\
%$\tau                  $   & $7.2$    \\
%$\text{b-tagging}       $  & $0.3$    \\
%$\text{Signal Theory}  $   & $1.5$   \\
%$\text{Other MC theory} $  & $2.7$    \\
%$t\bar{t} \text{theory}$  & $7.6$     \\\midrule
%$\text{Statistics}      $  & $11.2$   \\
%$\text{Total systematic}$  & $24.2$  \\ \midrule
%$\text{Total}  $           & $26.7$ \\
%\bottomrule\bottomrule
%\end{tabular}
%\label{tab:had_sys_impact}
%\end{table}

%\begin{table}[htbp]
%\caption{List of relative uncertainties of the signal strength from the combined fit in leptonic channel. The uncertainties are symmetrised for presentation and %grouped into  the categories described in the text. %The quadrature sum of the individual uncertainties is not equal to the total uncertainty due to correlations %introduced by the fit
%}
%\small
%\centering
%\begin{tabular}{cc} \toprule\toprule
%$\text{Uncertainty}        $ & $  \Delta\mu/\mu[\%]  $\\\midrule
%$\text{Fake modelling}     $ & $       2.1             $\\
%$\text{Instrument}         $ & $       0.7             $\\
%$\text{Jets}               $ & $       2.8             $\\
%$\text{Met}                $ & $       1.2             $\\
%$\text{Lepton}             $ & $       0.7             $\\
%$\tau                      $ & $       2.3             $\\
%$\text{b-tagging}          $ & $       4.8             $\\
%$\text{Signal Theory}      $ & $       6.9             $\\
%$\text{Other MC theory}    $ & $       2.1             $\\
%$t\bar{t} \text{theory}    $ & $       3.7             $\\\midrule
%$\text{Statistics}         $ & $       5.9             $\\
%$\text{Total systematic}   $ & $       12.9            $\\\midrule
%$\text{Total}              $ & $       14.2            $\\
%\bottomrule\bottomrule
%\end{tabular}
%\label{tab:lep_sys_impact}
%\end{table}