### 2022-06-30 sections/background_estimation.tex

%-------------------------------------------------------------------------------
\section{Background estimation}
\label{sec:background_model}
%-------------------------------------------------------------------------------

Most background processes are modelled using MC simulation.
After the event preselection, the main background is $\ttbar$ production, often in association with jets, denoted by $\ttbar$+jets in the following.
Small contributions arise from single-top-quark, $W/Z$+jets, multi-jet and diboson ($WW,WZ,ZZ$) production, as well as from the associated 
production of a vector boson $V$ ($V=W,Z$) or a Higgs boson and a $\ttbar$ pair ($\ttbar V$ and $\ttbar H$). All backgrounds 
with prompt leptons, i.e.\ those originating from the decay of a $W$ boson, a $Z$ boson, or a $\tau$-lepton,
are estimated using samples of simulated events and are initially normalised to their theoretical cross sections.
In the simulation, the top-quark and SM Higgs boson masses are set to $172.5~\gev$ and $125~\gev$, respectively,
and the Higgs boson is forced to decay via $H\to \tau\tau$ with a branching ratio calculated using \textsc{Hdecay}~\cite{Djouadi:1997yw}.  
Backgrounds with non-prompt light leptons (electron or muon), with photons or jets misidentified as electrons, or with jets misidentified as $\tau$-lepton candidates, 
generically referred to as fake leptons, are estimated using data-driven methods. 
%\texttt{fake} 
The background prediction is improved during the statistical analysis by performing a likelihood 
fit to data using several signal-depleted control regions as shown in Table~\ref{tab:srcr}.
The events with one light lepton, two same-sign $\had$, and one $b$-jet, are also selected, providing a validation region (VR) for the background estimation in the leptonic channels. 
%-------------------------------------------------------------------------------
%\subsection{Backgrounds with fake leptons}
%\label{sec:fakeleptons}
%-------------------------------------------------------------------------------
\subsection{Backgrounds with fake $\tau$-leptons}
\label{sec:faketaus}
The background with one or more fake $\tau$ candidates arises mainly from $\ttbar$ or
multi-jet production, depending on the search channel.
Studies based on simulation show that, for all the above processes, fake $\tau$ candidates primarily result from the
misidentification of light jets and $b$-quark jets.
It is also found that the fake rate decreases for all jet flavours as the $\tau$ candidate's $\pt$ increases.

In the leptonic channels, the events with a prompt electron or muon and fake $\tau$-leptons are modelled by calibrating the MC samples with scale factors (SF)
derived from the dedicated $t\bar t$
control regions (CRtt) using dileptonic decays of $t\bar t$ pairs and semileptonic decays of $t\bar t$ pairs with two
$b$-jets, as is summarised in Table~\ref{tab:srcr}. The control regions are defined similarly to the signal regions but with an additional $b$-jet or light lepton. 	
%the SM $t\bar t$ decay of dilepton events and semileptonically double-btagged lepton-jet events, summarized in Table~\ref{tab:srcr},
%aimed at fake taus from different origins.
There are four kinds of fake $\tau$-leptons that need to be calibrated: Type-1 fake $\tau$-leptons from hadronic $W$ boson decay ($\tau_{W}$) %$W$-jets ($\tau_{W}$)
with opposite-sign (OS) charge relative to the light lepton;
Type-2 $\tau_{W}$'s with same-sign (SS) charge relative to the light lepton; Type-3 fake $\tau$-leptons originating from $b$-hadron decays; and Type-4 fake $\tau$-leptons from light-flavour hadron decays.
%The control regions are defined similar to the signal regions but with an additional $b$-jet or light lepton.
%$t_{\ell}t_{\ell}1b\thad$, $t_{\ell}t_{\ell}2b\thad$, $t_{\ell}t_h2b\thad$-2jSS, $t_{\ell}t_h2b\thad$-2jOS, $t_{\ell}t_h2b\thad$-3jSS, and $t_{\ell}t_h2b\thad$-3jOS.
The dilepton regions ($t_{\ell}t_{\ell}1b\thad$ and $t_{\ell}t_{\ell}2b\thad$) are used to calibrate Type-3 and Type-4 fake $\tau$-leptons. The semileptonic
regions ($t_{\ell}t_h2b\thad$-2jOS and $t_{\ell}t_h2b\thad$-3jOS), where the $\thad$ and light lepton have opposite charges, are used to calibrate Type-1 fake $\tau$-leptons.
Similarly for Type-2, the semileptonic regions ($t_{\ell}t_h2b\thad$-2jSS and $t_{\ell}t_h2b\thad$-3jSS), where the $\thad$ and light lepton have same charges, are used.
A simultaneous fit to data is made to derive the scale factors for fake $\tau$-leptons in the MC samples. These consist of a total of 24 parameters
depending on four types of fake $\tau$-leptons, three $\pt$ bins, and two bins for 1- and 3-prong $\tau$ decays separately. The post-fit distributions of the
leading $\thad$ $\pt$ and measured $\tau$ scale factors are presented in Appendix~\ref{sec:tauFF_appendix}. 
The values of these scale factors range from 0.3 to 1.28 and are used to correct the MC-estimated fakes in the corresponding  signal regions with a single $b$-jet.
In the $t_{\ell}\thadhad$ channel, both $\tau$ candidates can be misidentified, so the calibration is applied to each $\tau$ candidate separately, following the same procedure used in the $\tlhad$ channel.
%using the lepton and fake tau charges, then the scale factors are multiplied together.
The central values of the scale factors vary according to their uncertainties in the final fit. A closure test is made for the fake-$\tau$ estimates, using the same procedure as
in the $t_{\ell}\thadhad$-SS validation region, and it shows good agreement between data and the background prediction. 
%In the control regions with single lepton, $\met > 20$GeV and at least 2 light jets and 2 b-tagged jets are required to ensure that QCD contribution is negligible.
%The calibration is done depending on different source of the fake taus and \pt. The calibration factor are derived in the dedicated
%control regions discussed in Section~\ref{sec:strategy_Htautau}.

In the hadronic channels, the contribution of fakes is estimated from data using the fake-factor (FF) method to derive the transfer functions from the background-enriched control
regions (CR) to the SR~\cite{ATLAS-CONF-2021-044}.
%by defining control regions (CR)
%enriched in misidentified
%tau candidates via loosened tau requirements with fake factors and partially from MC~\cite{ATLAS-CONF-2021-044}.
These CRs do not overlap with the main signal regions discussed in Section~\ref{sec:strategy_Htautau}.
The CR selection requirements are analogous to those used to define signal regions, except
that the subleading $\tau$ candidate is required to fail the `medium' $\tau$ identification while still passing a loose requirement.
The contribution of fakes with subleading $\tau$ candidates can be calculated by rescaling the templates for `loose' $\tau$ candidates in the CR
with the FFs.
The templates are produced by subtracting all MC background contributions with real subleading $\tau$-leptons from the data.
The FFs are computed as the ratio of misidentified $\thad$ candidates that either pass or fail the `medium' $\tau$ ID selection in regions enriched in events originating from 
$W$+jets processes~\cite{ATLAS-CONF-2021-044}. 
%the Data-MC ($\mathrm{real~tau}$) ratio of the yields of events in which the tau candidate is passing or failing the medium tau ID selection
%in the $W$+jets samples~\cite{ATLAS-CONF-2021-044}.
%passing to failing the medium tau ID in the $W$+jets control region.
%We have compared
FFs obtained from the $W$+jets events are compared with those from the same-sign $\thadhad$ control regions and the differences are treated as systematic uncertainties.

\subsection{Background with fake light leptons}
The background originating from non-prompt or misidentified light leptons is primarily from multi-jet production.
%Electrons and muons only appear in the leptonic channels where exactly one electron or muon and at least one $\had$ are required.
%Since the rate of $\had$ candidate is much larger than for electron or muon, in the majority of the fake lepton events, the $\had$ candidate is misidentified and
%the electron or muon is prompt based on MC studies.
%However the multi-jet background has large event rate and can contribute with both electron or muon and $\had$ faked, especially to $t_{\ell}\had$ where the fake-lepton dominates. and the jet multiplicity is low.
The contribution from these events is estimated with a data-driven method called ABCD.
The estimate is based on the numbers of `tight' and `loose' light leptons that passed and failed the PLIV cut in the low ($<25~\gev$) and high
($>25~\gev$) $\met$ regions. The number of
fake light leptons in the signal region is evaluated by scaling the number of `loose' leptons in the high $\met$ region by the ratio of `tight' leptons to `loose' leptons in the
low $\met$ region. The contributions of prompt leptons and calibrated fake $\tau$-leptons to the numbers of `tight' and `loose' leptons are subtracted using MC simulation.
A closure test is made for the background estimates in the signal-depleted low-BDT-score ($<-0.6$) regions, defined in Section~\ref{sec:tmva}.
The data are in good agreement with the background prediction in the
$t_{\ell}\had$ channels, while the fake light leptons are negligible in other leptonic channels.